---
title: "Analysis on species-specific pathways"
author: "Xiaoxi Dong"
output:
        html_document:
                toc: yes
editor_options: 
  chunk_output_type: console
---

# Number of pathway per species distribution
This analysis used pre-defined pathway-specific completeness cutoffs.

```{r, distributions,echo=FALSE, results='hide', message=FALSE}
d1 = load( "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/pathway.completeness.cutoff.info.RData")
d1
d2 = load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/pathway.species.pct_Mapped.RData")
d2
species.pathway.id = pathway.species.pct_Mapped
head(species.pathway.id[species.pathway.id[,"species"] == "hsa",])
min(species.pathway.id[species.pathway.id[,"species"] == "hsa","pct.mapped.species.pathway"])
range(species.pathway.id[species.pathway.id[,"species"] == "mmu","pct.mapped.species.pathway"])
range(species.pathway.id[species.pathway.id[,"species"] == "ptr","pct.mapped.species.pathway"])
mean(species.pathway.id[species.pathway.id[,"species"] == "mmu","pct.mapped.species.pathway"])
mean(species.pathway.id[species.pathway.id[,"species"] == "ptr","pct.mapped.species.pathway"])
median(species.pathway.id[species.pathway.id[,"species"] == "mmu","pct.mapped.species.pathway"])
median(species.pathway.id[species.pathway.id[,"species"] == "ptr","pct.mapped.species.pathway"])
length(species.pathway.id[species.pathway.id[,"species"] == "mmu","pct.mapped.species.pathway"])
length(species.pathway.id[species.pathway.id[,"species"] == "ptr","pct.mapped.species.pathway"])

d2
head(pathway.completeness.cutoff.info)
head(pathway.species.pct_Mapped)
```

```{r, prepare data,echo=FALSE, results='hide', message=FALSE}
library(SBGNview)
all.pathways = find.pathways(org="all")
dim(all.pathways)
head(all.pathways)
dim(pathway.species.pct_Mapped)
library(ggplot2)

#species name mapping
species.name.mapping = "C:/xiaoxi/work/input/id.mapping/kegg.related/organism.txt"
species.to.name = read.table(species.name.mapping,stringsAsFactors = F,sep="\t",header=F,quote="",comment.char = "")
colnames(species.to.name) = c("id","abr","full.name","path")
species.to.name = species.to.name[!duplicated(species.to.name$full.name),]
row.names(species.to.name) = species.to.name[,2]
head(species.to.name)

# species
species.ct = table(all.pathways$species)
head(species.ct)
species.ct = species.ct[order(species.ct,decreasing = T)]
species.ct[1:20]
class(names(species.ct))
species.ct = data.frame(species = names(species.ct),count = as.vector(species.ct),stringsAsFactors = F)
head(species.ct)
class(species.ct$species)
full.name = as.character(species.to.name[species.ct$species,"full.name"])
head(full.name)
head(species.ct$species)
species.ct$species = factor( full.name
                            ,levels = full.name
                            )
head(species.ct)
dim(species.ct)
```


Number of pathway per species distribution. Violin plot's width represent density.

```{r,number of pathways per species,echo=FALSE, results='hide', message=FALSE, fig.width=2}
all.count = species.ct$count
head(all.count)
all.count = data.frame(count = all.count, species = "all.species")
head(all.count)
p = ggplot(data=all.count
      ,aes(x=species,y=count)
      ,main = "Number of pathways found in each species: \n Showing top 20 species with largest number of pathways"
      )+ 
        # geom_boxplot(stat="boxplot" , fill="blue"# ,color="blue"
        #              ,outlier.size = 0.5
        #          ) +
        geom_violin(
                fill="blue"
        )+
      xlab("")+
      ylab("number of pathways")+ scale_y_log10() +scale_x_discrete(position = "bottom")
print(p)
```


## Top 20 species with largest number of pathways


```{r,number of pathways in species with most pathways,echo=FALSE, results='hide', message=FALSE}
# distribution of all species
species.ct.top = species.ct[1:20,]
species.ct.top
species.ct.top$species = factor(species.ct.top$species
                            ,levels = rev(species.ct.top$species)
                            )

p = ggplot(data=species.ct.top
      ,aes(x=species,y=count)
      ,main = "Number of pathways found in each species: \n Showing top 20 species with largest number of pathways"
      )+ 
        geom_bar(stat="identity" , fill="blue"# ,color="blue"
                 ) +
      xlab("species")+
      ylab("number of pathways")+
        coord_flip()+ scale_y_reverse(position = "right") +scale_x_discrete(position = "top")
print(p)

```

# Number of species per pathway distribution.  Violin plot's width represent density.


```{r,number of species per pathway,echo=FALSE, results='hide', message=FALSE, fig.width=2}

pathway.ct = table(all.pathways$pathway.id)
head(pathway.ct)
head(pathway.ct)
pathway.ct = pathway.ct[order(pathway.ct,decreasing = T)]
pathway.ct[1:20]
class(names(pathway.ct))
pathway.ct = data.frame(pathway = names(pathway.ct),count = as.vector(pathway.ct),stringsAsFactors = F)

all.count = pathway.ct$count
head(all.count)
all.count = data.frame(count = all.count, pathway = "all.pathway")
head(all.count)
p = ggplot(data=all.count
      ,aes(x=pathway,y=count)
      )+ 
        # geom_boxplot(stat="boxplot" , fill="blue"# ,color="blue"
        #              ,outlier.size = 0.5
        #          ) +
        geom_violin(
                fill="blue"
        )+
      xlab("")+
      ylab("number of species")+ scale_y_log10() +scale_x_discrete(position = "bottom")
print(p)
```

## Check some pathways distributed most widely or narrowly

```{r, check some pathways, echo=FALSE,message=FALSE}
data("pathway.info.pre.generated.files")
# head(pathway.info.pre.generated.files)
selected.pathways = pathway.ct
# head(pathway.ct)
selected.pathways = selected.pathways[!duplicated((selected.pathways[,"count"])),]
selected.pathways = selected.pathways[c(1:5,(nrow(selected.pathways)-5):nrow(selected.pathways)),]
# selected.pathways
pathway.to.name = pathway.info.pre.generated.files[,c("pathway.id","DISPLAY_NAME")]
colnames(pathway.to.name) = c("pathway","pathway.name")
# head(pathway.to.name)
# selected.pathways
selected.pathways = merge(selected.pathways,pathway.to.name)
selected.pathways = selected.pathways[order(selected.pathways$count),]
# head(selected.pathways)
selected.pathways[,3:2]
selected.pathways$pathway.name = factor(selected.pathways$pathway.name
                            ,levels = selected.pathways$pathway.name[order(selected.pathways$count)]
                            )

p = ggplot(data=selected.pathways
      ,aes(x=pathway.name,y=count)
      )+ 
        geom_bar(stat="identity" , fill="blue"# ,color="blue"
                 ) +
      xlab("pathway")+
      ylab("number of species where pathway is detected")+
        coord_flip()+scale_y_log10() +scale_x_discrete(position = "top") +
        ylim(0,max(selected.pathways$count))
print(p)
# selected.pathways
```
In these examples, the most narrowly distributed pathways are mostly signaling pathways. The most widely distributed pathways are the ones involved in basic metabolism processes.


## The species where the most narrowly distributed pathways are found:

```{r,most narrowly distributed pathways, echo=FALSE,message=FALSE}
sl = all.pathways[all.pathways[,"pathway.id"] %in% selected.pathways[1:5,"pathway"],c("DISPLAY_NAME","species")]
tapply(sl$species,as.factor(sl$DISPLAY_NAME)
       ,function(sp){
                full.name = as.character(species.to.name[sp,"full.name"])
        }
)
```

Check what happened to the most narrowly distributed pathway "mineralocorticoid biosynthesis: MetaCyc PWY66-382"
```{r,PWY66-382,echo=FALSE,message=FALSE}
narrow = species.pathway.id[species.pathway.id[,"pathway"] == "PWY66-382",]
narrow = narrow[order(narrow[,"pct.mapped.species.pathway"],decreasing=T),]
ggplot(narrow,aes(pct.mapped.species.pathway)) +geom_bar(position="stack") + xlab("pathway completeness: fraction") +ylab("number of pathways") + ggtitle("pathway completeness distribution: PWY66-382. This pathway has 2 KOs. \n Completeness 0.5 means one KO is detected in a species and\n the other KO (K07433) is missing.")
load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/mapping.each.pair/for.use.all/both/KO_pathway.id.RData")
ko.to.pathway.id = mapping.table
ko.to.pathway.id[,"pathway.id"] = as.character(ko.to.pathway.id[,"pathway.id"])
ko.to.pathway.id[,"KO"] = as.character(ko.to.pathway.id[,"KO"])
narrow.path.ko = with(ko.to.pathway.id,KO[pathway.id=="PWY66-382"])
```


It looks like this pathway has only two KOs. One of the KO (K07433) is missing in a lot of species.

```{r,message=FALSE}
read.table("http://rest.kegg.jp/link/genes/K07433")
```
K07433 was only detected in 6 species, according to KEGG ortholog definition. This is why the pathway is predicted to exist in only 5 species( the other KO:K00513 is missing in pps, so it is 5=6-1)

KEGG doesn't have this pathway.
In MetaCyc, it is annotated to exist in Mammalia:
https://biocyc.org/META/NEW-IMAGE?type=PATHWAY&object=PWY66-382
In reactome, this pathway is annotated in many other species, as distant as Saccharomyces cerevisiae
https://reactome.org/content/detail/R-HSA-193993
It doesn't make much sense that S.cerevisiae has this pathway because "Mineralocorticoids are a class of corticosteroids, which in turn are a class of steroid hormones. Mineralocorticoids are produced in the adrenal cortex and influence salt and water balances."

So species-specific pathway prediction is affected by ortholog prediction:
KEGG's BlastKOALA only uses sequence similarity but reactome's Compara fits similarity network of proteins to species tree. 
It looks like MetaCyc's species distribution (Mammalia) makes more sense. 

To solve this issue, maybe I can try different ortholog annotation databases/algorithms


```{r,Angiogenesis,eval=FALSE, echo=FALSE,message=FALSE}
library(SBGNview)
all.pathways.1 = find.pathways(pathway.completeness = 0, org="all")
head(all.pathways.1)
Angiogenesis = all.pathways.1[all.pathways.1[,"DISPLAY_NAME"] %in% "Angiogenesis" ,]
dim(Angiogenesis)
head(Angiogenesis)
library(ggplot2)
class(Angiogenesis[,"pct.mapped.species.pathway"])
head(Angiogenesis[,"pct.mapped.species.pathway"])
in.data = data.frame(completeness = as.vector(Angiogenesis[,"pct.mapped.species.pathway"])
                     ,pathway = "Angiogenesis"
                     )
p = ggplot(data= in.data
      ,aes(x=pathway,y=completeness)
      )+ 
        geom_violin() +
      xlab("")+
      ylab("completeness")+scale_x_discrete(position = "bottom") +
        geom_jitter()

print(p)
cutoff.to.f = cutoff.to.fs(Angiogenesis)
head(cutoff.to.f)
p = qplot(
        as.numeric(cutoff.to.f[,"cutoff"])
        ,as.numeric(cutoff.to.f[,"stat.F"])
        
)
plot(p)

```



# Distribution of F statistics.  Violin plot's width represent density.
```{r,distribution of F statistic,echo=FALSE, results='hide', message=FALSE, fig.width=2}
head(pathway.completeness.cutoff.info)
pathway.completeness.cutoff.info.anova = pathway.completeness.cutoff.info[!grepl("number", pathway.completeness.cutoff.info$ANOVA.F.stat),]
max(as.numeric(pathway.completeness.cutoff.info.anova$ANOVA.F.stat[!grepl("NA",pathway.completeness.cutoff.info.anova$ANOVA.F.stat)]))

all.count = as.numeric(pathway.completeness.cutoff.info.anova$ANOVA.F.stat)
head(all.count)
all.count = data.frame(count = all.count, pathway = "all.pathway")
head(all.count)
p = ggplot(data=all.count
      ,aes(x=pathway,y=count)
      )+ 
        # geom_boxplot(stat="boxplot" , fill="blue"# ,color="blue"
        #              ,outlier.size = 0.5
        #          ) +
        geom_violin(
                fill="blue"
        )+
      xlab("")+
      ylab("F statistics")+ scale_y_log10() +scale_x_discrete(position = "bottom")
print(p)
```



```{r,other distributions, distributions,echo=FALSE,eval=FALSE}
d1 = load( "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/pathway.completeness.cutoff.info.RData")
d1
d2 = load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/pathway.species.pct_Mapped.RData")
d2
head(pathway.completeness.cutoff.info)
head(pathway.species.pct_Mapped)
detach("package:SBGNview", unload=TRUE)
library(SBGNview)
all.pathways = find.pathways()
dim(all.pathways)
head(all.pathways)
max(all.pathways)
dim(pathway.species.pct_Mapped)
max(species.ct)
pathway.ct = table(all.pathways$pathway.id)
head(pathway.ct)
library(ggplot2)


#species name mapping
species.name.mapping = "C:/xiaoxi/work/input/id.mapping/kegg.related/organism.txt"
species.to.name = read.table(species.name.mapping,stringsAsFactors = F,sep="\t",header=F,quote="",comment.char = "")
colnames(species.to.name) = c("id","abr","full.name","path")
species.to.name = species.to.name[!duplicated(species.to.name$full.name),]
row.names(species.to.name) = species.to.name[,2]
head(species.to.name)

# species
species.ct = table(all.pathways$species)
head(species.ct)
species.ct = species.ct[order(species.ct,decreasing = T)]
species.ct[1:20]
class(names(species.ct))
# distribution of all species
species.ct = data.frame(species = names(species.ct),count = as.vector(species.ct),stringsAsFactors = F)
head(species.ct)
class(species.ct$species)
full.name = as.character(species.to.name[species.ct$species,"full.name"])
head(full.name)
head(species.ct$species)
species.ct$species = factor( full.name
                            ,levels = full.name
                            )
head(species.ct)
species.ct.top = species.ct[1:20,]
species.ct.top
species.ct.top$species
species.ct.top$species = factor(species.ct.top$species
                            ,levels = rev(species.ct.top$species)
                            )

class(species.ct[1:20,1])
p = ggplot(data=species.ct.top
      ,aes(x=species,y=count)
      ,main = "number of pathways each species has"
      )+ 
        geom_bar(stat="identity" , fill="blue"# ,color="blue"
                 ) +
      xlab("species")+
      ylab("number of pathways")+
        coord_flip()+ scale_y_reverse(position = "right") +scale_x_discrete(position = "top")
print(p)


# distribution of all species
p = qplot(species.ct
      ,geom="histogram"
      ,binwidth = 1
      ,main = "number of pathways each species has"
      ,ylab = "number of species"
      ,xlab = "number of pathways"
      ,xlim = c(0,200)
      )
print(p)


```


``` {r, eval = FALSE, echo=FALSE}
rmarkdown::render("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/distribution.species.pathways.Rmd")
```
