---
title: "Analysis on species-specific pathways"
author: "Xiaoxi Dong"
output:
        html_document:
                toc: yes
editor_options: 
  chunk_output_type: console
---

# Number of pathway per species distribution
This analysis used pre-defined pathway-specific completeness cutoffs.

```{r, distributions,echo=FALSE, results='hide', message=FALSE}
d1 = load( "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/pathway.completeness.cutoff.info.RData")
d1
d2 = load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/pathway.species.pct_Mapped.RData")
d2
head(pathway.completeness.cutoff.info)
head(pathway.species.pct_Mapped)
```

```{r, prepare data,echo=FALSE, results='hide', message=FALSE}
library(SBGNview)
all.pathways = find.pathways(org="all")
dim(all.pathways)
head(all.pathways)
dim(pathway.species.pct_Mapped)
library(ggplot2)

#species name mapping
species.name.mapping = "C:/xiaoxi/work/input/id.mapping/kegg.related/organism.txt"
species.to.name = read.table(species.name.mapping,stringsAsFactors = F,sep="\t",header=F,quote="",comment.char = "")
colnames(species.to.name) = c("id","abr","full.name","path")
species.to.name = species.to.name[!duplicated(species.to.name$full.name),]
row.names(species.to.name) = species.to.name[,2]
head(species.to.name)

# species
species.ct = table(all.pathways$species)
head(species.ct)
species.ct = species.ct[order(species.ct,decreasing = T)]
species.ct[1:20]
class(names(species.ct))
species.ct = data.frame(species = names(species.ct),count = as.vector(species.ct),stringsAsFactors = F)
head(species.ct)
class(species.ct$species)
full.name = as.character(species.to.name[species.ct$species,"full.name"])
head(full.name)
head(species.ct$species)
species.ct$species = factor( full.name
                            ,levels = full.name
                            )
head(species.ct)
dim(species.ct)
```


Number of pathway per species distribution. Violin plot's width represent density.

```{r,number of pathways per species,echo=FALSE, results='hide', message=FALSE, fig.width=2}
all.count = species.ct$count
head(all.count)
all.count = data.frame(count = all.count, species = "all.species")
head(all.count)
p = ggplot(data=all.count
      ,aes(x=species,y=count)
      ,main = "Number of pathways found in each species: \n Showing top 20 species with largest number of pathways"
      )+ 
        # geom_boxplot(stat="boxplot" , fill="blue"# ,color="blue"
        #              ,outlier.size = 0.5
        #          ) +
        geom_violin(
                fill="blue"
        )+
      xlab("")+
      ylab("number of pathways")+ scale_y_log10() +scale_x_discrete(position = "bottom")
print(p)
```


## Top 20 species with largest number of pathways


```{r,number of pathways in species with most pathways,echo=FALSE, results='hide', message=FALSE}
# distribution of all species
species.ct.top = species.ct[1:20,]
species.ct.top
species.ct.top$species = factor(species.ct.top$species
                            ,levels = rev(species.ct.top$species)
                            )

p = ggplot(data=species.ct.top
      ,aes(x=species,y=count)
      ,main = "Number of pathways found in each species: \n Showing top 20 species with largest number of pathways"
      )+ 
        geom_bar(stat="identity" , fill="blue"# ,color="blue"
                 ) +
      xlab("species")+
      ylab("number of pathways")+
        coord_flip()+ scale_y_reverse(position = "right") +scale_x_discrete(position = "top")
print(p)

```

# Number of species per pathway distribution.  Violin plot's width represent density.


```{r,number of species per pathway,echo=FALSE, results='hide', message=FALSE, fig.width=2}

pathway.ct = table(all.pathways$pathway.id)
head(pathway.ct)
head(pathway.ct)
pathway.ct = pathway.ct[order(pathway.ct,decreasing = T)]
pathway.ct[1:20]
class(names(pathway.ct))
pathway.ct = data.frame(pathway = names(pathway.ct),count = as.vector(pathway.ct),stringsAsFactors = F)

all.count = pathway.ct$count
head(all.count)
all.count = data.frame(count = all.count, pathway = "all.pathway")
head(all.count)
p = ggplot(data=all.count
      ,aes(x=pathway,y=count)
      )+ 
        # geom_boxplot(stat="boxplot" , fill="blue"# ,color="blue"
        #              ,outlier.size = 0.5
        #          ) +
        geom_violin(
                fill="blue"
        )+
      xlab("")+
      ylab("number of species")+ scale_y_log10() +scale_x_discrete(position = "bottom")
print(p)
```

## Check some pathways distributed most widely or narrowly

```{r, check some pathways, echo=FALSE,message=FALSE}
data("pathway.info.pre.generated.files")
# head(pathway.info.pre.generated.files)
selected.pathways = pathway.ct
# head(pathway.ct)
selected.pathways = selected.pathways[!duplicated((selected.pathways[,"count"])),]
selected.pathways = selected.pathways[c(1:5,(nrow(selected.pathways)-5):nrow(selected.pathways)),]
# selected.pathways
pathway.to.name = pathway.info.pre.generated.files[,c("pathway.id","DISPLAY_NAME")]
colnames(pathway.to.name) = c("pathway","pathway.name")
# head(pathway.to.name)
# selected.pathways
selected.pathways = merge(selected.pathways,pathway.to.name)
selected.pathways = selected.pathways[order(selected.pathways$count),]
# head(selected.pathways)
selected.pathways[,3:2]
selected.pathways$pathway.name = factor(selected.pathways$pathway.name
                            ,levels = selected.pathways$pathway.name[order(selected.pathways$count)]
                            )

p = ggplot(data=selected.pathways
      ,aes(x=pathway.name,y=count)
      )+ 
        geom_bar(stat="identity" , fill="blue"# ,color="blue"
                 ) +
      xlab("pathway")+
      ylab("number of species where pathway is detected")+
        coord_flip()+scale_y_log10() +scale_x_discrete(position = "top") +
        ylim(0,max(selected.pathways$count))
print(p)
```


## The species where the most narrowly distributed pathways are found:

```{r,most narrowly distributed pathways, echo=FALSE,message=FALSE}
sl = all.pathways[all.pathways[,"pathway.id"] %in% selected.pathways[1:5,"pathway"],c("DISPLAY_NAME","species")]
tapply(sl$species,as.factor(sl$DISPLAY_NAME)
       ,function(sp){
                full.name = as.character(species.to.name[sp,"full.name"])
        }
)
```

In these examples, the most narrowly distributed pathways are mostly signaling pathways. The most widely distributed pathways are the ones involved in basic metabolism processes.


```{r,Angiogenesis,eval=FALSE, echo=FALSE,message=FALSE}
library(SBGNview)
all.pathways.1 = find.pathways(pathway.completeness = 0, org="all")
head(all.pathways.1)
Angiogenesis = all.pathways.1[all.pathways.1[,"DISPLAY_NAME"] %in% "Angiogenesis" ,]
dim(Angiogenesis)
head(Angiogenesis)
library(ggplot2)
class(Angiogenesis[,"pct.mapped.species.pathway"])
head(Angiogenesis[,"pct.mapped.species.pathway"])
in.data = data.frame(completeness = as.vector(Angiogenesis[,"pct.mapped.species.pathway"])
                     ,pathway = "Angiogenesis"
                     )
p = ggplot(data= in.data
      ,aes(x=pathway,y=completeness)
      )+ 
        geom_violin() +
      xlab("")+
      ylab("completeness")+scale_x_discrete(position = "bottom") +
        geom_jitter()

print(p)
cutoff.to.f = cutoff.to.fs(Angiogenesis)
head(cutoff.to.f)
p = qplot(
        as.numeric(cutoff.to.f[,"cutoff"])
        ,as.numeric(cutoff.to.f[,"stat.F"])
        
)
plot(p)

```



# Distribution of F statistics.  Violin plot's width represent density.
```{r,distribution of F statistic,echo=FALSE, results='hide', message=FALSE, fig.width=2}
head(pathway.completeness.cutoff.info)
pathway.completeness.cutoff.info.anova = pathway.completeness.cutoff.info[!grepl("number", pathway.completeness.cutoff.info$ANOVA.F.stat),]
max(as.numeric(pathway.completeness.cutoff.info.anova$ANOVA.F.stat[!grepl("NA",pathway.completeness.cutoff.info.anova$ANOVA.F.stat)]))

all.count = as.numeric(pathway.completeness.cutoff.info.anova$ANOVA.F.stat)
head(all.count)
all.count = data.frame(count = all.count, pathway = "all.pathway")
head(all.count)
p = ggplot(data=all.count
      ,aes(x=pathway,y=count)
      )+ 
        # geom_boxplot(stat="boxplot" , fill="blue"# ,color="blue"
        #              ,outlier.size = 0.5
        #          ) +
        geom_violin(
                fill="blue"
        )+
      xlab("")+
      ylab("F statistics")+ scale_y_log10() +scale_x_discrete(position = "bottom")
print(p)
```



```{r,other distributions, distributions,echo=FALSE,eval=FALSE}
d1 = load( "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/pathway.completeness.cutoff.info.RData")
d1
d2 = load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/pathway.species.pct_Mapped.RData")
d2
head(pathway.completeness.cutoff.info)
head(pathway.species.pct_Mapped)
detach("package:SBGNview", unload=TRUE)
library(SBGNview)
all.pathways = find.pathways()
dim(all.pathways)
head(all.pathways)
max(all.pathways)
dim(pathway.species.pct_Mapped)
max(species.ct)
pathway.ct = table(all.pathways$pathway.id)
head(pathway.ct)
library(ggplot2)


#species name mapping
species.name.mapping = "C:/xiaoxi/work/input/id.mapping/kegg.related/organism.txt"
species.to.name = read.table(species.name.mapping,stringsAsFactors = F,sep="\t",header=F,quote="",comment.char = "")
colnames(species.to.name) = c("id","abr","full.name","path")
species.to.name = species.to.name[!duplicated(species.to.name$full.name),]
row.names(species.to.name) = species.to.name[,2]
head(species.to.name)

# species
species.ct = table(all.pathways$species)
head(species.ct)
species.ct = species.ct[order(species.ct,decreasing = T)]
species.ct[1:20]
class(names(species.ct))
# distribution of all species
species.ct = data.frame(species = names(species.ct),count = as.vector(species.ct),stringsAsFactors = F)
head(species.ct)
class(species.ct$species)
full.name = as.character(species.to.name[species.ct$species,"full.name"])
head(full.name)
head(species.ct$species)
species.ct$species = factor( full.name
                            ,levels = full.name
                            )
head(species.ct)
species.ct.top = species.ct[1:20,]
species.ct.top
species.ct.top$species
species.ct.top$species = factor(species.ct.top$species
                            ,levels = rev(species.ct.top$species)
                            )

class(species.ct[1:20,1])
p = ggplot(data=species.ct.top
      ,aes(x=species,y=count)
      ,main = "number of pathways each species has"
      )+ 
        geom_bar(stat="identity" , fill="blue"# ,color="blue"
                 ) +
      xlab("species")+
      ylab("number of pathways")+
        coord_flip()+ scale_y_reverse(position = "right") +scale_x_discrete(position = "top")
print(p)


# distribution of all species
p = qplot(species.ct
      ,geom="histogram"
      ,binwidth = 1
      ,main = "number of pathways each species has"
      ,ylab = "number of species"
      ,xlab = "number of pathways"
      ,xlim = c(0,200)
      )
print(p)


```


``` {r, eval = FALSE, echo=FALSE}
rmarkdown::render("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/distribution.species.pathways.Rmd")
```
