---
title: "PathwayCommons id mapping"
author: "Xiaoxi Dong"
date: "October 12, 2018"
output:
        html_document:
                toc: yes
editor_options: 
  chunk_output_type: console
---
# Id mapping procedures


## proteins
### pathwayCommons 
#### pathwayCommons to human uniprot

script: 

```{r , echo = TRUE, message = FALSE, warning = FALSE}
source("./pathway.commons_SBGN.id.to.uniprot.chebi.using.biopax.R")
biopax.folder = "C:/xiaoxi/work/papers/SBGNview/package/data.package/pathway.commons/raw.data/v9"
if(!file.exists(paste(biopax.folder,"/small.molecules.mapping.list.rds",sep=""))){
    print("need to generate mapping file")
    biopax.files.to.id.mapping(biopax.folder)
}else{
    print("mapping file generated")}
```

```{r , echo = TRUE, message = FALSE, warning = FALSE}
biopax.folder = "C:/xiaoxi/work/papers/SBGNview/package/data.package/pathway.commons/raw.data/v9"
protein.mapping =  readRDS(paste(biopax.folder,"/proteins.mapping.list.rds",sep=""))
names(protein.mapping)
protein.mapping$duplicated.ids
nrow(protein.mapping$mappings.no.database.id)
head(protein.mapping$mappings.database.id.name)
pathwayCommons_UNIPROT_human = protein.mapping$mappings.database.id.name[,2:1]
colnames(pathwayCommons_UNIPROT_human) = c("UNIPROT","pathwayCommons")
head(pathwayCommons_UNIPROT_human)
```

#### map other species to pathwayCommons UNIPROT

##### Reactome id mapping for different species

Reactome has pathway information for many species, but they only provides SBGN files for homo sapiens. 
A file can map all species' uniprot ids to gene names:
https://reactome.org/download/current/UniProt2Reactome_PE_Pathway.txt
Maybe I can use this and specify the SBGN id to be "label"
But there is no documentation saying that the label is gene name.

###### How did Reactome generate pathway information for other species
The annotation for other species could be mostly come from IEA.
Count evidence code.
For each species in reactome database, count how many genes are annotated under each evidence code
``` {r, echo=T}
evidence.to.species.file = "C://xiaoxi/work/input/Reactome.web/UniProt2Reactome_PE_All_Levels.evidence.to.species.txt"
evidence.to.species = read.table(evidence.to.species.file,sep="\t",header=F,comment.char = "")
count.table = t(table(evidence.to.species))
print(count.table)
```
http://wiki.geneontology.org/index.php/Inferred_from_Electronic_Annotation_(IEA)

The IEA evidence code is used for annotations that are computationally, or automatically, assigned to gene products without further manual, curator review. IEA annotations are derived from two main pipelines: 1) manually constructed mappings between external classification systems and GO terms, and 2) automatic transfer of annotation to orthologous gene products.

IEA annotations must use a reference from the GO Consortium's collection of GO references to indicate the method whereby the IEA annotation was made. If a suitable reference is not available, curators may request a new GO reference where it will be given a GO_REF ID, e.g. GO_REF:0000002, that will be used as the reference in Gene Association Files (GAFs).
        
http://wiki.geneontology.org/index.php/Traceable_Author_Statement_(TAS)
The TAS evidence code covers author statements that are attributed to a cited reference that is the source of the original evidence (e,g, experimental results, sequence comparison, etc.).

###### How to generate SBGN map for other species
One route is:
other species' uniprot --> gene name --> human uniprot --> pathwayCommons id --> SBGN map

Generate mapping:
uniprot, gene name, species
```{bash}
cat C://xiaoxi/work/input/Reactome.web/UniProt2Reactome_PE_All_Levels.txt |cut -f1,3,8 >C://xiaoxi/work/input/Reactome.web/UniProt2Reactome_PE_All_Levels.uniprot.geneName.species.txt
```
```{r}

mapping = read.table("C://xiaoxi/work/input/Reactome.web/UniProt2Reactome_PE_All_Levels.uniprot.geneName.species.txt",sep="\t",stringsAsFactors = F,quote='')
# mapping = read.table("C://xiaoxi/work/input/Reactome.web/UniProt2Reactome_PE_All_Levels.uniprot.geneName.species.txt",sep="\t",stringsAsFactors = F)
mapping = unique(mapping)
dim(mapping)
mapping.no.compartment = mapping
head(mapping)
gene.name.no.compartment= gsub(" \\[.+","",mapping.no.compartment[,2])
head(gene.name.no.compartment)
mapping.no.compartment[,2] = gene.name.no.compartment
head(mapping.no.compartment)
mapping = mapping.no.compartment


mapping.human = mapping[mapping[,3]=="Homo sapiens",]
colnames(mapping.human) = c("uniprot","gene.name","species")
mapping.non.human = mapping[mapping[,3]!="Homo sapiens",]
colnames(mapping.non.human) = c("uniprot","gene.name","species")
head(mapping.human)
head(mapping.non.human)
mapping.nonHuman.to.human = merge(mapping.non.human,mapping.human,by="gene.name",suffixes = c(".non.human",".human"))
head(mapping.nonHuman.to.human)
dim(mapping.nonHuman.to.human)
mapping.nonHuman.to.human = mapping.nonHuman.to.human[,c("uniprot.non.human","species.non.human","uniprot.human")]
mapping.nonHuman.to.human = unique(mapping.nonHuman.to.human)
head(mapping.nonHuman.to.human)
dim(mapping.nonHuman.to.human)

write.table( mapping.nonHuman.to.human
            ,"./id.mapping/pathway.commons/reactome.other.species/non.human.uniprot.to.human.uniprot.txt"
            )


```
###### merge through human uniprot
```{r}
mapping.nonHuman.to.human = read.table("C:/xiaoxi/work/papers/SBGNview/package/id.mapping/id.mapping/pathway.commons/reactome.other.species/non.human.uniprot.to.human.uniprot.txt",
                                       stringsAsFactors = F,
                                       header=T,
                                       sep=" "
            )
head(mapping.nonHuman.to.human)
pathwayCommons_UNIPROT_human = cbind(pathwayCommons_UNIPROT_human,"Homo sapiens")
head(pathwayCommons_UNIPROT_human)
non.human.uniprot.to.pathwayCommons = merge(mapping.nonHuman.to.human,pathwayCommons_UNIPROT_human
                                            ,by.x = "uniprot.human"
                                            ,by.y="UNIPROT"
                                            )
dim(non.human.uniprot.to.pathwayCommons)
head(non.human.uniprot.to.pathwayCommons)
colnames(non.human.uniprot.to.pathwayCommons)[2:3] = c("uniprot","species")
head(non.human.uniprot.to.pathwayCommons)
```
merge human mapping and other species mapping
```{r}
uniprot.to.pathwayCommons.all.species = pathwayCommons_UNIPROT_human
colnames(uniprot.to.pathwayCommons.all.species)[3] = "species"
colnames(uniprot.to.pathwayCommons.all.species)[1] = "uniprot"
head(uniprot.to.pathwayCommons.all.species)
head(non.human.uniprot.to.pathwayCommons)
uniprot.to.pathwayCommons.all.species = rbind(uniprot.to.pathwayCommons.all.species
                                              ,non.human.uniprot.to.pathwayCommons[,c("uniprot","pathwayCommons","species")]
                                              )
colnames(uniprot.to.pathwayCommons.all.species)[1] = "UNIPROT"
head(uniprot.to.pathwayCommons.all.species)
tail(uniprot.to.pathwayCommons.all.species)
table(uniprot.to.pathwayCommons.all.species$species)
uniprot.to.pathwayCommons.all.species.list = list()
uniprot.to.pathwayCommons.all.species.list$gene = list()
uniprot.to.pathwayCommons.all.species.list$gene = uniprot.to.pathwayCommons.all.species 
save(uniprot.to.pathwayCommons.all.species.list,file="C:/xiaoxi/work/papers/SBGNview/package/id.mapping/id.mapping/pathway.commons/uniprot.to.pathwayCommons.all.species.list.RData")


```
###### Generate entrez to pathwayCommons using all species uniprot to pathwayCommons

```{r}
source("./all.ids.to.pathwayCommons.metacyc.R")
all.gene.id.types.to.pathwayCommons(uniprot.to.pathwayCommons.species.matrix = uniprot.to.pathwayCommons.all.species,gene.to.pathwayCommons.each.pair.RData.folder = "../data/id.mapping//mapping.each.pair/for.use.all/") 

```
##### other species uniprot to pathwayCommons through KEGG ortholog
###### other species uniprot to human uniprot in kegg
human uniprot to human entrez through kegg

```{r}
        
download.from.kegg = function(ids,url,output.file,size.slice){
        all.species.entrez = paste(ids,collapse = "+")
        kegg.q.all.species.entrez.to.all.species.uniprot = paste(url,all.species.entrez,sep="")
        substr(kegg.q.all.species.entrez.to.all.species.uniprot ,1,50)
        # print(kegg.q.all.species.entrez.to.all.species.uniprot)
        # stop()
        if.found = try(download.file(kegg.q.all.species.entrez.to.all.species.uniprot ,destfile = output.file,quiet = T,mode = "a"))
        if(class(if.found) == "try-error"){
                n.genes = length(ids)
                slices = seq(1,n.genes,by=size.slice)
                for(j in 1:length(slices)){
                        if(j == length(slices)){
                                end.index = n.genes
                        }else{
                                end.index = slices[j+1] -1
                        }
                        slice.genes = slices[j]:end.index
                        all.species.entrez = paste(ids[slice.genes],collapse = "+")
                        cat("downloading to ", output.file,"\n",paste(min(slice.genes),max(slice.genes),sep="-"),"out of",n.genes,"genes\n")
                        kegg.q.all.species.entrez.to.all.species.uniprot = paste(url,all.species.entrez,sep="")
                        substr(kegg.q.all.species.entrez.to.all.species.uniprot ,1,50)
                        if.downloaded = try(download.file(kegg.q.all.species.entrez.to.all.species.uniprot ,destfile = output.file,quiet = T,mode = "a"))
                        # if("up:A1L0T0" %in% ids[slice.genes]){
                        #         try(download.file(kegg.q.all.species.entrez.to.all.species.uniprot ,destfile = paste(output.file,".A1L0T0.tsv",sep=""),quiet = T,mode = "a"))
                        #         print(kegg.q.all.species.entrez.to.all.species.uniprot)
                        #         stop("found A1L0T0")
                        # }
                        if(class(if.downloaded) == "try-error"){
                                download.from.kegg(ids[slice.genes],url,output.file,(size.slice-1))
                        }
                }
        }
}


biopax.folder = "C:/xiaoxi/work/papers/SBGNview/package/data.package/pathway.commons/raw.data/v9"
protein.mapping =  readRDS(paste(biopax.folder,"/proteins.mapping.list.rds",sep=""))
names(protein.mapping)
protein.mapping$duplicated.ids
nrow(protein.mapping$mappings.no.database.id)
head(protein.mapping$mappings.database.id.name)
pathwayCommons_UNIPROT_human = protein.mapping$mappings.database.id.name[,2:1]
colnames(pathwayCommons_UNIPROT_human) = c("UNIPROT","pathwayCommons")
head(pathwayCommons_UNIPROT_human)
head(pathwayCommons_UNIPROT_human)
dim(pathwayCommons_UNIPROT_human)
human.uniprot.to.entrez.file = "kegg.human.uniprot.to.entrez.tsv"
human.entrez.to.ko.file = "kegg.human.entrez.to.ko.tsv"
all.species.entrez.to.ko.file = "kegg.ko.to.all.species.entrez.tsv"
kegg.humanUniprot.to.NonHumanUniprot.orthologs.file = "C:/xiaoxi/work/input/id.mapping/kegg/kegg.humanUniprot.to.NonHumanUniprot.orthologs.tsv"
# file.remove(human.uniprot.to.entrez.file,human.entrez.to.ko.file,all.species.entrez.to.uniprot.file,kegg.humanUniprot.to.NonHumanUniprot.orthologs.file)
all.uniprots = unique(pathwayCommons_UNIPROT_human[,"UNIPROT"])
n.pathwayCommons.uniprot = length(all.uniprots)
n.pathwayCommons.uniprot



# human uniprot to genes
current.uniprot = all.uniprots
output.folder = "C:/xiaoxi/work/input/id.mapping/kegg.ortholog.100.download.slice"
dir.create(output.folder)
output.file = paste(output.folder,"/",human.uniprot.to.entrez.file,sep="")
download.from.kegg (paste("up:",current.uniprot,sep=""),"http://rest.kegg.jp/conv/hsa/",output.file,size.slice = 100)
kegg.q.uniprot.to.entrez.human = try(read.table(output.file,header = F,stringsAsFactors = F,sep="\t",quote=""))
colnames(kegg.q.uniprot.to.entrez.human) =c("uniprot","entrez")
head(kegg.q.uniprot.to.entrez.human)

# entrez to ko
output.file = paste(output.folder,"/",human.entrez.to.ko.file,sep="")
download.from.kegg (unique(kegg.q.uniprot.to.entrez.human[,"entrez"]),"http://rest.kegg.jp/link/ko/",output.file,size.slice = 100)
kegg.human.entrez.to.ko = try(read.table(paste(output.folder,"/",human.entrez.to.ko.file,sep=""),header = F,stringsAsFactors = F,sep="\t",quote=""))
colnames(kegg.human.entrez.to.ko) =c("entrez.human","ko")
head(kegg.human.entrez.to.ko)

# ko to all species gene
output.file = paste(output.folder,"/",all.species.entrez.to.ko.file,sep="")
download.from.kegg (unique(kegg.human.entrez.to.ko[,"ko"]),"http://rest.kegg.jp/link/genes/",output.file,size.slice = 20)
kegg.ko.to.all.species.entrez = try(read.table(output.file,header = F,stringsAsFactors = F,sep="\t",quote=""))
colnames(kegg.ko.to.all.species.entrez ) =c("ko","all.species.entrez")
head(kegg.ko.to.all.species.entrez )
dim(kegg.ko.to.all.species.entrez )
table(kegg.ko.to.all.species.entrez[,"ko"] )
        
        # In the first round of run, only the first 100 of every 200 genes in a slice was downloaded
        finished.kos = read.table("C:/xiaoxi/work/input/id.mapping/kegg.species.ortholog.first.100.per.200/kegg.ko.to.all.species.entrez.tsv",header=F,sep="\t",stringsAsFactors=F)
        head(finished.kos)
        dim(finished.kos)
        finished.kos = unique(finished.kos[,1])
        head(finished.kos)
        length(finished.kos)
        unfinished.kos = setdiff(unique(kegg.ko.to.all.species.entrez[,"ko"]),finished.kos)
        length(unfinished.kos)
        kegg.ko.to.all.species.entrez.backup = kegg.ko.to.all.species.entrez
        dim(kegg.ko.to.all.species.entrez.backup)
        kegg.ko.to.all.species.entrez = kegg.ko.to.all.species.entrez[kegg.ko.to.all.species.entrez[,"ko"] %in% unfinished.kos,]
        dim(kegg.ko.to.all.species.entrez)
        
# all species entrez to all species uniprot
dim(kegg.ko.to.all.species.entrez)
uniprot.to.ko.file = paste(output.folder, "/uniprot.to.ko.tsv",sep="")
by(
        kegg.ko.to.all.species.entrez
        ,as.factor(kegg.ko.to.all.species.entrez[,"ko"])
        ,function(all.species.entrez.df){
                ko = all.species.entrez.df[1,"ko"]
                ko = gsub("ko:","",ko)
                ko.uniprot.file =  paste(output.folder,ko,".tsv",sep="")
                
                output.file = ko.uniprot.file
                if(file.exists(output.file)){
                        print(ko)
                        return()
                }
                download.from.kegg (unique(all.species.entrez.df[,"all.species.entrez"]),"http://rest.kegg.jp/conv/uniprot/",output.file,size.slice = 100)
                
                all.species.entrez.to.uniprot = read.table(ko.uniprot.file,header = F,stringsAsFactors = F,sep="\t",quote="")
                colnames(all.species.entrez.to.uniprot) =c("gene","uniprot")
                all.species.entrez.to.uniprot[,"uniprot"] = gsub("up:","",all.species.entrez.to.uniprot[,"uniprot"])
                species = do.call(rbind,strsplit(all.species.entrez.to.uniprot[,"gene"],":"))[,1]
                uniprot.to.ko = all.species.entrez.to.uniprot[,c("uniprot","gene")]
                uniprot.to.ko = cbind(uniprot.to.ko,ko)
                write.table(uniprot.to.ko,uniprot.to.ko.file,row.names = F,col.names =F,sep="\t",quote=F,append = T)
                # stop()
        }
)

uniprot.to.ko.file
# map other species uniprot to human uniprot
uniprot.to.ko = read.table(uniprot.to.ko.file,sep="\t",stringsAsFactors = F,header = F)
colnames(uniprot.to.ko) = c("uniprot","gene","ko")
dim(uniprot.to.ko)
uniprot.to.ko = unique(uniprot.to.ko)
dim(uniprot.to.ko)
head(uniprot.to.ko)
species = do.call(rbind,strsplit(uniprot.to.ko[,"gene"],":"))[,1]
uniprot.to.ko[,2] = species
colnames(uniprot.to.ko) = c("uniprot","species","ko")
head(uniprot.to.ko)
human.uniprot = uniprot.to.ko[uniprot.to.ko[,"species"] == "hsa",]
non.human.uniprot = uniprot.to.ko[uniprot.to.ko[,"species"] != "hsa",]

dim(human.uniprot)
dim(non.human.uniprot)
head(human.uniprot)
head(non.human.uniprot)
length(unique(uniprot.to.ko[,"ko"]))
length(unique(uniprot.to.ko[,"species"]))
head(unique(uniprot.to.ko[,"species"]))

non.human.to.human = merge(human.uniprot,non.human.uniprot,by="ko",suffixes = c(".human",".non.human"),all=F)

head(non.human.to.human)
tail(non.human.to.human)
dim(non.human.to.human)

# other species uniprot to pathwayCommons by human uniprot
head(non.human.to.human)
head(pathwayCommons_UNIPROT_human)
non.human.uniprot.to.pathwayCommons = merge(non.human.to.human[,c("uniprot.human","uniprot.non.human","species.non.human")],pathwayCommons_UNIPROT_human
                                            ,by.x = "uniprot.human"
                                            ,by.y="UNIPROT"
                                            ,all = F
                                            )
dim(non.human.uniprot.to.pathwayCommons)
head(non.human.uniprot.to.pathwayCommons)
colnames(non.human.uniprot.to.pathwayCommons)[2:3] = c("uniprot","species")
head(non.human.uniprot.to.pathwayCommons)
head(unique(non.human.uniprot.to.pathwayCommons[,5]))
# combine other species to pathwayCommons and human uniprot to pathwayCommons
uniprot.to.pathwayCommons.all.species = pathwayCommons_UNIPROT_human
class(uniprot.to.pathwayCommons.all.species)
uniprot.to.pathwayCommons.all.species = as.data.frame(uniprot.to.pathwayCommons.all.species,stringsAsFactors  = F)
uniprot.to.pathwayCommons.all.species$species = "hsa"
colnames(uniprot.to.pathwayCommons.all.species)[3] = "species"
colnames(uniprot.to.pathwayCommons.all.species)[1] = "uniprot"
head(uniprot.to.pathwayCommons.all.species)
head(non.human.uniprot.to.pathwayCommons)
uniprot.to.pathwayCommons.all.species = rbind(uniprot.to.pathwayCommons.all.species
                                              ,non.human.uniprot.to.pathwayCommons[,c("uniprot","pathwayCommons","species")]
                                              )
colnames(uniprot.to.pathwayCommons.all.species)[1] = "UNIPROT"
head(uniprot.to.pathwayCommons.all.species)
tail(uniprot.to.pathwayCommons.all.species)
table(uniprot.to.pathwayCommons.all.species$species)
uniprot.to.pathwayCommons.all.species.kegg.list = list()
uniprot.to.pathwayCommons.all.species.kegg.list$gene = list()
uniprot.to.pathwayCommons.all.species.kegg.list$gene = uniprot.to.pathwayCommons.all.species 
save(uniprot.to.pathwayCommons.all.species.kegg.list,file="C:/xiaoxi/work/papers/SBGNview/package/id.mapping/id.mapping/pathway.commons/uniprot.to.pathwayCommons.all.species.kegg.list.RData")

# generate usable id mapping list :pathwayCommons to uniprot: kegg ortholog definition
load("C:/xiaoxi/work/papers/SBGNview/package/id.mapping/id.mapping/pathway.commons/uniprot.to.pathwayCommons.all.species.kegg.list.RData")
pathwayCommons_UNIPROT = uniprot.to.pathwayCommons.all.species.kegg.list$gene
head(pathwayCommons_UNIPROT)
tail(pathwayCommons_UNIPROT)
# head(pathwayCommons_UNIPROT[pathwayCommons_UNIPROT[,"species"]=="Homo sapiens",])
# pathwayCommons_UNIPROT[,"species"] = as.character(pathwayCommons_UNIPROT[,"species"])
# pathwayCommons_UNIPROT[pathwayCommons_UNIPROT[,"species"]=="Homo sapiens","species"] = "hsa"
head(pathwayCommons_UNIPROT)
mapping.list = list()
mapping.list$gene = list()
mapping.list$gene$pathwayCommons_UNIPROT = pathwayCommons_UNIPROT
head(mapping.list$gene$pathwayCommons_UNIPROT)
dim(mapping.list$gene$pathwayCommons_UNIPROT)
save(mapping.list,file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/mapping.each.pair/for.use.all/both/pathwayCommons_UNIPROT.RData")
# generate usable id mapping list :pathwayCommons to ko
head(uniprot.to.ko)
length(unique(uniprot.to.ko$ko))
human.uniprot.to.ko = uniprot.to.ko[grepl("hsa",uniprot.to.ko[,2]),]
human.uniprot.to.ko = human.uniprot.to.ko[,c(1,3)]
colnames(human.uniprot.to.ko) = c("UNIPROT","KO")
head(human.uniprot.to.ko)
length(unique(human.uniprot.to.ko$KO))
head(unique(human.uniprot.to.ko$KO))
dim(human.uniprot.to.ko)
head(pathwayCommons_UNIPROT_human)
KO.to.pathwayCommons = merge(human.uniprot.to.ko
                             ,pathwayCommons_UNIPROT_human
                             ,all=F
                             )
head(KO.to.pathwayCommons)
dim(KO.to.pathwayCommons)
length(unique(KO.to.pathwayCommons$KO))
type.pair.name = paste(sort(c("KO","pathwayCommons")),collapse="_")
type.pair.name
KO_pathwayCommons = KO.to.pathwayCommons[,2:3]
head(KO_pathwayCommons)
dim(KO_pathwayCommons)
mapping.list = list()
mapping.list$gene = list()
mapping.list$gene$KO_pathwayCommons = KO_pathwayCommons
head(mapping.list$gene$KO_pathwayCommons)
dim(mapping.list$gene$KO_pathwayCommons)
save(mapping.list,file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/mapping.each.pair/for.use.all/both/KO_pathwayCommons.RData")

```

Check which pathway the ko are from
```{r}
# ko to all species gene
all.species.entrez.to.ko.file = "kegg.ko.to.all.species.entrez.tsv"
output.folder = "C:/xiaoxi/work/input/id.mapping/kegg"
output.file = paste(output.folder,"/",all.species.entrez.to.ko.file,sep="")
# download.from.kegg (unique(kegg.human.entrez.to.ko[,"ko"]),"http://rest.kegg.jp/link/genes/",output.file,size.slice = 20)
kegg.ko.to.all.species.entrez = try(read.table(output.file,header = F,stringsAsFactors = F,sep="\t",quote=""))
head(kegg.ko.to.all.species.entrez)
colnames(kegg.ko.to.all.species.entrez ) =c("ko","all.species.entrez")
ko.species = do.call(rbind,strsplit(kegg.ko.to.all.species.entrez[,"all.species.entrez"],":"))[,1]
head(ko.species)
head(kegg.ko.to.all.species.entrez )
dim(kegg.ko.to.all.species.entrez )

kegg.species.count = tapply(
        ko.species
        ,as.factor(kegg.ko.to.all.species.entrez[,"ko"])
        ,function(species){
                length(unique(species))
        }
)
kegg.species.count = kegg.species.count[order(kegg.species.count,decreasing = T)]
head(kegg.species.count)


```
###### KO to entrezid
```{r, ko to uniprot}
eg.to.ko = read.table("C:/xiaoxi/work/input/id.mapping/kegg.related/kegg.ortholog.100.download.slice/kegg.ko.to.all.species.entrez.tsv",sep="\t",stringsAsFactors = F,quote="")
head(eg.to.ko)
eg.to.ko[,1] = gsub("ko:","",eg.to.ko[,1])
eg.to.ko[,3] = do.call(rbind,strsplit(eg.to.ko[,2],":"))[,1]
eg.to.ko[,2] = do.call(rbind,strsplit(eg.to.ko[,2],":"))[,2]
head(eg.to.ko)
dim(eg.to.ko)
class(eg.to.ko[,1])
class(eg.to.ko[,2])
class(eg.to.ko[,3])
colnames(eg.to.ko) = c("KO","ENTREZID","species")
tail(eg.to.ko)
head(eg.to.ko)
mapping.table = eg.to.ko
save(mapping.table,file="C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/KO_ENTREZID.RData")
```
###### KO to uniprot
```{r, ko to uniprot}
write.table(uniprot.to.ko,"C:/xiaoxi/work/input/id.mapping/kegg.ortholog.100.download.slice/uniprot.to.ko.tsv",row.names=F,sep="\t")
ko.to.uniprot = read.table("C:/xiaoxi/work/input/id.mapping/kegg.related/kegg.ortholog.100.download.slice/uniprot.to.ko.tsv",header=T,stringsAsFactors = F,sep="\t")
head(ko.to.uniprot)
ko.to.uniprot = ko.to.uniprot[,c(1,3)]
colnames(ko.to.uniprot) = c("UNIPROT","KO")
head(ko.to.uniprot)
mapping.list = list()
mapping.list$gene = list()
mapping.list$gene$UNIPROT_KO = ko.to.uniprot
save(mapping.list,file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/mapping.each.pair/for.use.all/both/UNIPROT_KO.RData")

```
###### KO to metacyc
```{r, ko to meacyc}
head(ko.to.uniprot)
dim(ko.to.uniprot)
length(unique(ko.to.uniprot$KO))
load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/mapping.each.pair/for.use.all/both/metacyc.SBGN_UNIPROT.RData")
uniprot.to.metacyc = mapping.list$gene$metacyc.SBGN_UNIPROT
head(uniprot.to.metacyc)
ko.to.metacyc = merge(ko.to.uniprot,uniprot.to.metacyc[,1:2],all=F)
head(ko.to.metacyc)
ko.to.metacyc = ko.to.metacyc[,c("KO","metacyc.SBGN")]
head(ko.to.metacyc)
dim(ko.to.metacyc)
ko.to.metacyc = ko.to.metacyc[!grepl("NA:@:NA",ko.to.metacyc[,"metacyc.SBGN"]),]
dim(ko.to.metacyc)
mapping.list = list()
mapping.list$gene = list()
mapping.list$gene$KO_metacyc.SBGN = ko.to.metacyc
save(mapping.list,file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/mapping.each.pair/for.use.all/both/KO_metacyc.SBGN.RData")
```


<!--         C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\pathway.commons\pathway.commons_SBGN.id.to.uniprot.chebi.using.biopax.R -->
<!-- input: -->

<!--         pathwayCommons biopax files in folder: /media/sf_work/input/id.mapping/pathway.commons.10.biopax -->
<!-- output: -->

<!-- ###################################################################### -->
        <!-- C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\pathway.commons\SBGNview.uniprot.chebi.to.SBGN.id.mapping.RData -->
        <!-- In this data, the uniprot ids are all from human -->
<!-- ###################################################################### -->

1.1 Other species uniprot to human uniprot, to pathwayCommons' reactome

script:

       C:\xiaoxi\work\papers\SBGNview\package\id.mapping\Reactome.Rmd
       
input: 

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\pathway.commons\reactome.other.species\UniProt2Reactome_PE_All_Levels.txt
        
output: 


        C:\xiaoxi\work\papers\SBGNview\package\SBGN-ML.files\data\id.mapping\mapping.each.pair\gene\pathwayCommons\derived.from.raw\non.human.uniprot.to.human.uniprot_pathwayCommons-reactome.txt
        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping/pathway.commons\uniprot.to.pathwayCommons.all.species.list.RData



2. uniprot to entrez

Use pathview

3. pathwayCommons to entrez

script: 

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\all.ids.to.pathwayCommons.metacyc.R
        entrenz.to.pathwayCommons()

input:

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping/id.mapping/pathway.commons/uniprot.to.pathwayCommons.all.species.list.RData

output: 

<!-- ###################################################################### -->

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\pathway.commons.and.metacyc\entrezid.to.metacyc.pathwayCommons.with.species.info.RData
<!-- ###################################################################### -->
        
        
        
        
### plant reactome
#### find all pathway names
1. cat /c/xiaoxi/work/input/plant.reactome/*.txt |cut -f2|sor
t|uniq>/c/xiaoxi/work/input/plant.reactome/all.pathways.txt
2. cat /c/xiaoxi/work/input/plant.reactome/all.pathways.txt |
cut -d'-' -f3|sort|uniq >/c/xiaoxi/work/input/plant.reactome
/all.pathways.number.txt
#### download pathways
```{r}
all.pathways = read.table("c:/xiaoxi/work/input/plant.reactome/all.pathways.number.txt",sep="\t",stringsAsFactors = F,header=F)
head(all.pathways)
dim(all.pathways)
link = "http://plantreactome.gramene.org/ReactomeRESTfulAPI/RESTfulWS/biopaxExporter/Level3/"
for (i in 1:nrow(all.pathways)){
        adress = paste(link,all.pathways[i,1],sep="")
        adress
        download.file(adress,destfile = paste("c:/xiaoxi/work/input/plant.reactome/downloaded/",all.pathways[i,1],".biopax"),sep="")
}
```
#### change biopax to sbgn
```{r}

```
        
### metacyc 

1. metacyc to uniprot/uniref

1.1 Use perl to change metacyc raw files to pairwise mapping tables:

script: 
        
        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\metacyc/Metacyc.raw.file.to.mapping.tables.pl

input:

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\metacyc\raw.data
        
output 

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\metacyc\mapping.tables
        
1.2 download metacyc biopax files

script :

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\metacyc\biopax.downloaded\download.metacyc.pl

result:

        C:\xiaoxi\work\input\2018.10.3.linux\metacyc\biopax.level3
        
1.3 from metacyc biopax files, extract mapping : metacyc sbgn/biopax vertex/glyph id --> metacyc id

script:

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\metacyc/metacyc.SBGN.id.to.uniprot.uniref.chebi.R  ::  generate.biopax.id.to.chebi.or.metacyc()
        
result: 

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\metacyc\mapping.tables\proteins.to.metacyc.mapping.all.RData
        
1.4 use pairwise mapping tables to map metacyc SBGN ids to uniprot/uniref

script:

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\metacyc/metacyc.SBGN.id.to.uniprot.uniref.chebi.R
        uniref.to.biopax.id()
        uniprot.to.biopax.id()
        
input: 

        MetaCyc: mapping tables downloaded from metacyc. They send a link to download after registering. The files are in the rar file. I selected the files needed for mapping and put them in this table:
        
        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\metacyc\mapping.tables
        Uniref: Used mapping table from humann2 package: metacyc_reactions_level4ec_only.uniref

result: 

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\metacyc\merged.tables
<!-- ###################################################################### -->

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\metacyc\merged.tables\uniref.pathway.sbgnBiopaxProtein.RData
        uniprot.pathway.sbgnBiopaxProtein.RData
<!-- ###################################################################### -->

check examples:
```{r}
load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/UNIPROT_pathway.id.RData")
mapping = mapping.table[mapping.table[,2] %in% c("R-HSA-877300","TOLUENE-DEG-CATECHOL-PWY"，"SMP02313","P00001"),]
mapping = mapping[order(mapping[,2],mapping[,1]),]
mapping = unique(mapping)
mapping = mapping[mapping[,2] == "TOLUENE-DEG-CATECHOL-PWY",]
mapping

load("C:/xiaoxi/work/papers/SBGNview/package/id.mapping/id.mapping/metacyc/merged.tables/metacycProtein.sbgnBiopaxProtein.pathway.enzymReaction.Reaction.uniprot.mapping.RData")
head(metacycProtein.sbgnBiopaxProtein.pathway.enzymReaction.Reaction.uniprot.mapping)
merged = subset(metacycProtein.sbgnBiopaxProtein.pathway.enzymReaction.Reaction.uniprot.mapping
                ,biopax.file == "TOLUENE-DEG-CATECHOL-PWY")
subset(merged,Uniprot == "P05091")

load("C:/xiaoxi/work/papers/SBGNview/package/id.mapping/id.mapping/pathway.commons.and.metacyc/entrezid.to.metacyc.pathwayCommons.with.species.info.RData")
head(entrezid.to.metacyc.pathwayCommons.with.species.info)
entrez.metacyc = entrezid.to.metacyc.pathwayCommons.with.species.info$ENTREZID_metacyc.SBGN
head(entrez.metacyc)
entrez.to.metacyc = entrez.metacyc[grepl(pattern = "TOLUENE-DEG-CATECHOL-PWY"
                                         ,x = entrez.metacyc[,"metacyc.SBGN"])
                                   ,
                                   ]
head(entrez.to.metacyc)


load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/ENTREZID_pathway.id.RData")

mapping = mapping.table[mapping.table[,2] %in% c("TOLUENE-DEG-CATECHOL-PWY"),]
mapping


```

        
1.5  metacyc sbgn ids --> entrez

        
script: 

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\all.ids.to.pathwayCommons.metacyc.R
        entrenz.to.metacyc.gene()

input:

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\metacyc\merged.tables/uniprot.pathway.sbgnBiopaxProtein.RData

output: 

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\pathway.commons.and.metacyc\entrezid.to.metacyc.pathwayCommons.with.species.info.RData
        
### text/symbol/gene name to molecule id: pathwayCommons metacyc
```{r}
library(SBGNview)
data("pathways.info")
source("C:/xiaoxi/work/papers/SBGNview/package/SBGNview/R/parsing.utilities.R")
source("C:/xiaoxi/work/papers/SBGNview/package/SBGNview/R/plot.utilities.R")
source("C:/xiaoxi/work/papers/SBGNview/package/SBGNview/R/SBGN.to.SVG.R")
is.pathwayCommons = pathways.info[,"database"] %in% c("pathwayCommons")
pathwayCommons.ids = pathways.info[is.pathwayCommons ,"pathway.id"]
length(pathwayCommons.ids)
node.list.pathwayCommons = sbgn.nodes(input.sbgn = c( 
                                        # pathwayCommons.ids
                                        pathwayCommons.ids[1:6]
                                        # "R-HSA-909733"
                                        )
                       ,sbgn.folder = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/SBGN/pathwayCommons"
                               )
head(node.list.pathwayCommons[[5]])
length(node.list.pathwayCommons)

is.MetaCyc = pathways.info[,"database"] %in% c("MetaCyc")
MetaCyc.ids = pathways.info[is.MetaCyc ,"pathway.id"]
length(MetaCyc.ids)
node.list.MetaCyc = sbgn.nodes(input.sbgn = c( 
                                        MetaCyc.ids
                                        # pathwayCommons.ids[1:3]
                                        # "R-HSA-909733"
                                        )
                       ,sbgn.folder = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/SBGN/MetaCyc"
                               )
head(node.list.MetaCyc[[5]])
length(node.list.MetaCyc)
node.list = c(node.list.pathwayCommons,node.list.MetaCyc)
length(node.list)

```
### text/symbol/gene name to molecule id: pathwayCommons metacyc
```{r}

label.to.glyph.id.all = matrix(ncol=2,nrow=0)
for (i in 1:length(node.list)){
# for (i in 1:3){
        print(i)
        label.to.class = node.list[[i]][,c("class","label","glyph.id")]
        print("generating dataframe")
        label.to.class = as.data.frame(label.to.class,stringsAsFactors=F)
        class(label.to.class)
        head(label.to.class)
        print("extracting macromolecule")
        macromolecule.labels = subset(label.to.class,class=="macromolecule",select="label")
        glyph.ids = subset(label.to.class,class=="macromolecule",select="glyph.id")
        head(macromolecule.labels)
        if(nrow(macromolecule.labels) == 0){
                next()
        }
        label.to.glyph.id = cbind(macromolecule.labels,glyph.ids)
        label.to.glyph.id = unique(label.to.glyph.id)
        label.to.glyph.id
        head(label.to.glyph.id)
        colnames(label.to.glyph.id) = c("SYMBOL","molecule.id")
        head(label.to.glyph.id)
        label.to.glyph.id.all = rbind(label.to.glyph.id.all,label.to.glyph.id)
}
label.to.glyph.id.all = unique(label.to.glyph.id.all)
label.to.glyph.id.all.no.species = label.to.glyph.id.all
label.to.glyph.id.all = cbind(label.to.glyph.id.all
                              ,"species")

dim(label.to.glyph.id.all)
head(label.to.glyph.id.all)
label.to.glyph.id.all[,2] = as.character(label.to.glyph.id.all[,2])
label.to.glyph.id.all[,3] = as.character(label.to.glyph.id.all[,3])
class(label.to.glyph.id.all[,1])
class(label.to.glyph.id.all[,2])
class(label.to.glyph.id.all[,3])
colnames(label.to.glyph.id.all)[3] = "species"
tail(label.to.glyph.id.all)

aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/pathwayCommons_SYMBOL.without.label.RData")
aa
mapping.table = mapping.list$gene[[1]]
dim(mapping.table)
head(mapping.table)
mapping.table[,1] = as.character(mapping.table[,1])
mapping.table[,2] = as.character(mapping.table[,2])
mapping.table[,3] = as.character(mapping.table[,3])
class(mapping.table[,1])
class(mapping.table[,2])
class(mapping.table[,3])
colnames(label.to.glyph.id.all) = colnames(mapping.table)
head(mapping.table)
tail(label.to.glyph.id.all)
mapping.table = rbind(mapping.table,label.to.glyph.id.all)
head(mapping.table)
dim(mapping.table)
mapping.table = unique(mapping.table)
dim(mapping.table)
tail(mapping.table)
head(mapping.table)
class(mapping.table[,1])
mapping.list$gene[[1]] = mapping.table
tail(mapping.list$gene[[1]])
save(mapping.list,file="C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/pathwayCommons_SYMBOL.RData")
aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/pathwayCommons_SYMBOL.RData")
aa
head(mapping.list$gene$pathwayCommons_SYMBOL)
tail(mapping.list$gene$pathwayCommons_SYMBOL)
dim(mapping.list$gene$pathwayCommons_SYMBOL)
aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/pathwayCommons_SYMBOL.without.label.RData")
dim(mapping.list$gene$pathwayCommons_SYMBOL)
write.table(mapping.list$gene$pathwayCommons_SYMBOL,"C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/pathwayCommons_SYMBOL.without.label.RData.tsv",row.names=F,sep="\t")
aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/pathwayCommons_SYMBOL.RData")
dim(mapping.list$gene$pathwayCommons_SYMBOL)
head(mapping.list$gene$pathwayCommons_SYMBOL)
write.table(mapping.list$gene$pathwayCommons_SYMBOL,"C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/pathwayCommons_SYMBOL.RData.tsv",row.names=F,sep="\t")

aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/metacyc.SBGN_SYMBOL.without.label.RData")
aa
mapping.table = mapping.list$gene[[1]]
dim(mapping.table)
head(mapping.table)
mapping.table[,1] = as.character(mapping.table[,1])
mapping.table[,2] = as.character(mapping.table[,2])
mapping.table[,3] = as.character(mapping.table[,3])
class(mapping.table[,1])
class(mapping.table[,2])
class(mapping.table[,3])
class(label.to.glyph.id.all[,1])
class(label.to.glyph.id.all[,2])
class(label.to.glyph.id.all[,3])
colnames(label.to.glyph.id.all) = colnames(mapping.table)
head(mapping.table)
tail(label.to.glyph.id.all)
mapping.table = rbind(mapping.table,label.to.glyph.id.all)
head(mapping.table)
dim(mapping.table)
mapping.table = unique(mapping.table)
dim(mapping.table)
tail(mapping.table)
head(mapping.table)
class(mapping.table[,1])
mapping.list$gene[[1]] = mapping.table
tail(mapping.list$gene[[1]])
save(mapping.list,file="C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/metacyc.SBGN_SYMBOL.RData")
aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/metacyc.SBGN_SYMBOL.RData")
aa
head(mapping.list$gene$metacyc.SBGN_SYMBOL)
tail(mapping.list$gene$metacyc.SBGN_SYMBOL)
dim(mapping.list$gene$metacyc.SBGN_SYMBOL)
aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/metacyc.SBGN_SYMBOL.without.label.RData")
aa
head(mapping.list$gene$metacyc.SBGN_SYMBOL)
tail(mapping.list$gene$metacyc.SBGN_SYMBOL)
dim(mapping.list$gene$metacyc.SBGN_SYMBOL)
```







### ko to glyph.ids update using label to glyph.ids
```{r}
aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/KO_metacyc.SBGN.RData")
aa
ko.to.metacyc = mapping.list$gene$KO_metacyc.SBGN
colnames(ko.to.metacyc) = c("KO","glyph.id")
head(ko.to.metacyc)

aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/KO_pathwayCommons.RData")
aa
ko.to.pathwayCommons = mapping.list$gene$KO_pathwayCommons
colnames(ko.to.pathwayCommons) = c("KO","glyph.id")
head(ko.to.pathwayCommons)

ko.to.glyph.ids = rbind(ko.to.metacyc
                        ,ko.to.pathwayCommons)
head(ko.to.glyph.ids)
class(ko.to.glyph.ids[,1])
class(ko.to.glyph.ids[,2])


# aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/pathwayCommons_SYMBOL.issue.id.RData")
aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/pathwayCommons_SYMBOL.RData")
aa
pathwayCommons.to.symbol = mapping.list$gene$pathwayCommons_SYMBOL
pathwayCommons.to.symbol[grepl("Protein53117",pathwayCommons.to.symbol[,"pathwayCommons"]),]
# pathwayCommons.to.symbol[,"pathwayCommons"] = gsub("\\.\\.\\."
#                                                    ,":@:"
#                                                    ,pathwayCommons.to.symbol[,"pathwayCommons"]
#                                                    )
# pathwayCommons.to.symbol[grepl("Protein53117",pathwayCommons.to.symbol[,"pathwayCommons"]),]
# mapping.list$gene$pathwayCommons_SYMBOL = pathwayCommons.to.symbol
# save(mapping.list, file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/pathwayCommons_SYMBOL.RData")
head(pathwayCommons.to.symbol)


# load metacyc
# aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/metacyc.SBGN_SYMBOL.issue.id.RData")
aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/metacyc.SBGN_SYMBOL.RData")
metacyc.to.symbol = mapping.list$gene$metacyc.SBGN_SYMBOL
head(metacyc.to.symbol)
tail(metacyc.to.symbol)
# metacyc.to.symbol[grepl("Protein53117",metacyc.to.symbol[,"metacyc.SBGN"]),]
# metacyc.to.symbol[,"metacyc.SBGN"] = gsub("\\.\\.\\."
#                                                    ,":@:"
#                                                    ,metacyc.to.symbol[,"metacyc.SBGN"]
#                                                    )
# metacyc.to.symbol[grepl("Protein53117",metacyc.to.symbol[,"metacyc.SBGN"]),]
# mapping.list$gene$metacyc.SBGN_SYMBOL = metacyc.to.symbol
# save(mapping.list, file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/metacyc.SBGN_SYMBOL.RData")
head(metacyc.to.symbol)

colnames(pathwayCommons.to.symbol)[2] ="glyph.id"
colnames(metacyc.to.symbol)[2] ="glyph.id"
symbol.to.glyph.id = rbind(pathwayCommons.to.symbol
                           ,metacyc.to.symbol)
symbol.to.glyph.id = unique(symbol.to.glyph.id)
head(symbol.to.glyph.id)
dim(symbol.to.glyph.id)

all.symbols = unique(symbol.to.glyph.id[,"SYMBOL"])
head(all.symbols)
output.type = "KO"
input.type = "SYMBOL"
species = "hsa"
cpd.or.gene = "gene"
limit.to.ids = all.symbols
length(all.symbols)

```
#### KO to symbol and gene name using sbgnview/pathview mapping
```{r}

library(SBGNview)
source("C:/xiaoxi/work/papers/SBGNview/package/SBGNview/R/parsing.utilities.R")
source("C:/xiaoxi/work/papers/SBGNview/package/SBGNview/R/plot.utilities.R")
source("C:/xiaoxi/work/papers/SBGNview/package/SBGNview/R/SBGN.to.SVG.R")
symbol.to.ko = load.mapping.file(
        output.type = "KO"
        ,input.type = "SYMBOL"
        ,species = "hsa"
        ,cpd.or.gene = "gene"
        ,limit.to.ids = toupper(all.symbols)
)
ko.to.symbol = symbol.to.ko$gene$KO_SYMBOL
dim(ko.to.symbol)
head(ko.to.symbol)

GENENAME.to.ko = load.mapping.file(
        output.type = "KO"
        ,input.type = "GENENAME"
        ,species = "hsa"
        ,cpd.or.gene = "gene"
        # ,limit.to.ids = toupper(all.symbols)
        ,limit.to.ids = all.symbols
)
ko.to.GENENAME = GENENAME.to.ko$gene$GENENAME_KO
dim(ko.to.GENENAME)
head(ko.to.GENENAME)

colnames(ko.to.GENENAME)[1] = "SYMBOL"
head(ko.to.GENENAME)
head(ko.to.symbol)
dim(ko.to.GENENAME)
dim(ko.to.symbol)
ko.to.symbol = rbind(ko.to.symbol
                     ,ko.to.GENENAME)
dim(ko.to.symbol)

```
#### merge to get final table
```{r}

###########
head(symbol.to.glyph.id)
symbol.to.glyph.id[grepl("Protein53117",symbol.to.glyph.id[,"glyph.id"]),]
tail(symbol.to.glyph.id)
class(symbol.to.glyph.id[,2])
head(ko.to.symbol)
ko.to.glyph.id.from.symbol = merge(ko.to.symbol,symbol.to.glyph.id,all=F)
ko.to.glyph.id.from.symbol = ko.to.glyph.id.from.symbol[,c("KO","glyph.id")]
ko.to.glyph.id.from.symbol = unique(ko.to.glyph.id.from.symbol)
head(ko.to.glyph.id.from.symbol)
dim(ko.to.glyph.id.from.symbol)

mapping.list = list()
mapping.list$gene = list()
colnames(ko.to.glyph.id.from.symbol)[2] = "metacyc.SBGN"
mapping.list$gene[["KO_metacyc.SBGN"]] = ko.to.glyph.id.from.symbol

head(mapping.list$gene$KO_metacyc.SBGN)
save(mapping.list,file="C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/KO_metacyc.SBGN.RData")



mapping.list = list()
mapping.list$gene = list()
colnames(ko.to.glyph.id.from.symbol)[2] = "pathwayCommons"
mapping.list$gene[["KO_pathwayCommons"]] = ko.to.glyph.id.from.symbol
head(mapping.list$gene$KO_pathwayCommons)
class(mapping.list$gene$KO_pathwayCommons[,1])
class(mapping.list$gene$KO_pathwayCommons[,2])
save(mapping.list,file="C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/KO_pathwayCommons.RData")


```
### ko to pathway.id
```{r}
pathway.to.glyph.id.all = matrix(ncol=2,nrow=0)
for (i in 1:length(node.list)){
# for (i in 1:3){
        print(i)
        pathway = names(node.list[i])
        pathway
        label.to.class = node.list[[i]][,c("class","label","glyph.id")]
        print("generating dataframe")
        label.to.class = as.data.frame(label.to.class,stringsAsFactors=F)
        class(label.to.class)
        head(label.to.class)
        print("extracting macromolecule")
        macromolecule.labels = subset(label.to.class,class=="macromolecule",select="label")
        glyph.ids = subset(label.to.class,class=="macromolecule",select="glyph.id")
        
        head(macromolecule.labels)
        if(nrow(macromolecule.labels) == 0){
                next()
        }
        glyph.ids
        pathway.to.glyph.id = cbind(glyph.ids,pathway)
        pathway.to.glyph.id = unique(pathway.to.glyph.id)
        pathway.to.glyph.id
        head(pathway.to.glyph.id)
        colnames(pathway.to.glyph.id) = c("SYMBOL","molecule.id")
        head(pathway.to.glyph.id)
        pathway.to.glyph.id.all = rbind(pathway.to.glyph.id.all,pathway.to.glyph.id)
}
pathway.to.glyph.id.all = unique(pathway.to.glyph.id.all)
head(pathway.to.glyph.id.all)
pathway.to.glyph.id.all[pathway.to.glyph.id.all[,"metacyc.SBGN"] == "Protein_7e80572b1f17bcfea96f8380de4ff8a5_Complex_af8cc7e6370ef36100e75db9fb86daf0",]

aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/pathwayCommons_pathway.id.RData")
aa
head(pathwayCommons_pathway.id)
dim(pathwayCommons_pathway.id)
colnames(pathway.to.glyph.id.all) = c("pathwayCommons","pathway.id")
head(pathwayCommons_pathway.id)
head(pathway.to.glyph.id.all)
pathwayCommons_pathway.id = rbind(pathwayCommons_pathway.id
                                  ,pathway.to.glyph.id.all)
dim(pathwayCommons_pathway.id)
pathwayCommons_pathway.id = unique(pathwayCommons_pathway.id)
dim(pathwayCommons_pathway.id)
mapping.table = pathwayCommons_pathway.id
head(mapping.table)
tail(mapping.table)
save(mapping.table, file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/pathwayCommons_pathway.id.RData")

aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/metacyc.SBGN_pathway.id.without.complex.member.RData")
aa
head(mapping.table)
dim(mapping.table)
colnames(pathway.to.glyph.id.all) = c("metacyc.SBGN","pathway.id")
head(mapping.table)
head(pathway.to.glyph.id.all)
tail(pathway.to.glyph.id.all)
mapping.table = rbind(mapping.table
                                  ,pathway.to.glyph.id.all)
dim(mapping.table)
mapping.table = unique(mapping.table)
dim(mapping.table)
head(mapping.table)
tail(mapping.table)
head(pathwayCommons_pathway.id)
save(mapping.table, file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/metacyc.SBGN_pathway.id.RData")
metacyc.pathway.to.metacyc.glyph = mapping.table
head(metacyc.pathway.to.metacyc.glyph)
dim(metacyc.pathway.to.metacyc.glyph)


```

```{r}
pathway.id.to.glyphs = metacyc.pathway.to.metacyc.glyph
colnames(pathway.id.to.glyphs)[1] = "pathwayCommons"
pathway.id.to.glyphs = rbind(pathway.id.to.glyphs,pathwayCommons_pathway.id)
pathway.id.to.glyphs = unique(pathway.id.to.glyphs)
colnames(pathway.id.to.glyphs)[1] = "glyph.id"
head(pathway.id.to.glyphs)
tail(pathway.id.to.glyphs)
dim(pathway.id.to.glyphs)
dim(mapping.table)
dim(pathwayCommons_pathway.id)

aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/KO_metacyc.SBGN.RData")
aa
ko.to.metacyc = mapping.list$gene$KO_metacyc.SBGN
colnames(ko.to.metacyc) = c("KO","glyph.id")
head(ko.to.metacyc)
aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/KO_pathwayCommons.RData")
aa
ko.to.pathwayCommons = mapping.list$gene$KO_pathwayCommons
colnames(ko.to.pathwayCommons) = c("KO","glyph.id")
head(ko.to.pathwayCommons)
ko.to.glyph.ids = rbind(ko.to.metacyc
                        ,ko.to.pathwayCommons)
ko.to.glyph.ids = unique(ko.to.glyph.ids)
head(ko.to.glyph.ids)
class(ko.to.glyph.ids[,1])
class(ko.to.glyph.ids[,2])
ko.to.glyph.ids[ko.to.glyph.ids[,"glyph.id"] == "Protein_7e80572b1f17bcfea96f8380de4ff8a5_Complex_af8cc7e6370ef36100e75db9fb86daf0",]
head(pathway.id.to.glyphs)
ko.to.pathway.id = merge(ko.to.glyph.ids
                         ,pathway.id.to.glyphs)
dim(ko.to.pathway.id)
ko.to.pathway.id = ko.to.pathway.id[,c("KO","pathway.id")]
head(ko.to.pathway.id)
ko.to.pathway.id = unique(ko.to.pathway.id)
dim(ko.to.pathway.id)
ko.to.pathway.id[ko.to.pathway.id[,"pathway.id"] == "R-HSA-909733" & ko.to.pathway.id[,"KO"] == "K11221",]

aa=load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/KO_pathway.id.RData")
aa
mapping.table = unique(mapping.table)
dim(mapping.table)
head(mapping.table)
dim(mapping.table[mapping.table[,"pathway.id" ]== "R-HSA-909733",])
dim(ko.to.pathway.id[ko.to.pathway.id[,"pathway.id" ]== "R-HSA-909733",])
mapping.table.all = rbind(mapping.table,ko.to.pathway.id)
mapping.table.all = unique(mapping.table.all)

dim(mapping.table)
dim(mapping.table.all)
dim(ko.to.pathway.id)
dim(mapping.table.all[mapping.table.all[,"pathway.id" ]== "R-HSA-909733",])
mapping.table = mapping.table.all
save(mapping.table,file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/KO_pathway.id.RData")



```


### update entrez to pathway, adding complex members

```{r}

library(SBGNview)
aa = load( "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/KO_ENTREZID.RData")
aa
head(mapping.table)
all.entrez = unique(mapping.table[,"ENTREZID"])
head(all.entrez)
length(all.entrez)


    source("C:/xiaoxi/work/papers/SBGNview/package/SBGNview/R/parsing.utilities.R")
    source("C:/xiaoxi/work/papers/SBGNview/package/SBGNview/R/plot.utilities.R")
    source("C:/xiaoxi/work/papers/SBGNview/package/SBGNview/R/SBGN.to.SVG.R")
entrez.to.symbol = load.mapping.file(
                        output.type = "SYMBOL"
                        ,input.type = "ENTREZID"
                        ,cpd.or.gene = "gene"
                        ,species = "hsa"
                        ,limit.to.ids = all.entrez
)
entrez.to.symbol = entrez.to.symbol[[1]][[1]]
head(entrez.to.symbol)
dim(entrez.to.symbol)

aa = load( "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/SYMBOL_pathway.id.RData")
aa
head(mapping.table)
head(entrez.to.symbol)
mapping.table[mapping.table[,"pathway.id" ]== "R-HSA-909733",]
entrez.to.pathway = merge(mapping.table,entrez.to.symbol)
entrez.to.pathway = entrez.to.pathway[,c("ENTREZID","pathway.id")]
head(entrez.to.pathway)
entrez.to.pathway[entrez.to.pathway[,"pathway.id" ]== "R-HSA-909733" & entrez.to.pathway[,"ENTREZID"] == "6773",]


# aa = load( "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/ENTREZID_pathway.id.RData")
aa = load( "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/ENTREZID_pathway.id.no.complex.members.RData")
aa
head(mapping.table)
head(entrez.to.pathway)
dim(mapping.table)
dim(entrez.to.pathway)
mapping.table[mapping.table[,"pathway.id" ]== "R-HSA-909733" & mapping.table[,"ENTREZID"] == "7297",]
length(unique(mapping.table[mapping.table[,"pathway.id" ]== "R-HSA-909733" ,"ENTREZID"]))

mapping.table = rbind(mapping.table,entrez.to.pathway)
dim(mapping.table)
mapping.table = unique(mapping.table)
dim(mapping.table)
length(unique(mapping.table[mapping.table[,"pathway.id" ]== "R-HSA-909733" ,"ENTREZID"]))
save(mapping.table, file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/ENTREZID_pathway.id.RData")



```

### text/symbol/gene name to pathway id: pathwayCommons metacyc
```{r}
label.to.pathway.all = matrix(ncol=2,nrow=0)
for (i in 1:length(node.list)){
        print(i)
        pathway = names(node.list)[i]
        pathway
        label.to.class = node.list[[i]][,c("class","label")]
        print("generating dataframe")
        label.to.class = as.data.frame(label.to.class,stringsAsFactors=F)
        class(label.to.class)
        head(label.to.class)
        print("extracting macromolecule")
        macromolecule.labels = subset(label.to.class,class=="macromolecule",select="label")
        if(nrow(macromolecule.labels) == 0){
                next()
        }
        label.to.pathway = cbind(macromolecule.labels,pathway)
        label.to.pathway = unique(label.to.pathway)
        label.to.pathway
        head(label.to.pathway)
        colnames(label.to.pathway) = c("SYMBOL","pathway.id")
        label.to.pathway.all = rbind(label.to.pathway.all,label.to.pathway)
}
label.to.pathway.all = unique(label.to.pathway.all)
dim(label.to.pathway.all)
head(label.to.pathway.all)
# add to symbol-pathway id mapping
load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/SYMBOL_pathway.id.without.node.label.RData")
dim(mapping.table)
mapping.table = rbind(mapping.table,label.to.pathway.all)
head(mapping.table)
dim(mapping.table)
mapping.table = unique(mapping.table)
dim(mapping.table)
tail(mapping.table)
head(mapping.table)
class(mapping.table[,1])
save(mapping.table,file="C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/SYMBOL_pathway.id.RData")

colnames(label.to.pathway.all)[1] = "GENENAME"
dim(label.to.pathway.all)
head(label.to.pathway.all)
# add to symbol-pathway id mapping
load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/GENENAME_pathway.id.without.node.label.RData")
dim(mapping.table)
head(mapping.table)
tail(mapping.table)
mapping.table = rbind(mapping.table,label.to.pathway.all)
tail(mapping.table)
dim(mapping.table)
mapping.table = unique(mapping.table)
dim(mapping.table)
tail(mapping.table)
head(mapping.table)
class(mapping.table[,1])
save(mapping.table,file="C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/GENENAME_pathway.id.RData")
```
## species specific gene ids to glyph id, pathway id
```{r}
library(SBGNview)
source("C:/xiaoxi/work/papers/SBGNview/package/SBGNview/R/parsing.utilities.R")
source("C:/xiaoxi/work/papers/SBGNview/package/SBGNview/R/plot.utilities.R")
source("C:/xiaoxi/work/papers/SBGNview/package/SBGNview/R/SBGN.to.SVG.R")
data(mapped.ids)
class(mapped.ids)
mapped.ids
all.genes = mapped.ids[mapped.ids[,"cpd.or.gene"]=="gene","ID.type.1"]
all.genes = unique(all.genes)
all.genes
all.genes.select = all.genes[c(4,5,6,10,12)]

library(SBGNview)
# all.species = c("hsa","mmu","ath")
all.species = c("mmu")
all.genes.select = "KO"
output.types = c("ENSEMBL")
i=1;j=1;k=1
for(i in 1:length(all.genes.select)){
       gene.type = all.genes.select[i]
       for(j in 1:length(output.types)){
               output.type = output.types[j]
               for(k in 1:length(all.species)){
                       species = all.species[k]
                       cat("\n\n\n",gene.type,"\n")
                       cat("  ",output.type,"\n")
                       cat("    ",species,"\n")
                        mapping.list = load.mapping.table(
                                output.type = output.type
                                ,input.type = gene.type
                                ,cpd.or.gene = "gene"
                                ,species = species
                        )
                        print(head(mapping.list[[1]][[1]]))
                        type.pair.name = paste(sort(c(gene.type,output.type),method = "radix",decreasing = TRUE),collapse="_") # R CMD check will give different sort result if you didn't specify "method":  the default sort method depends on the locale of your system. And R CMD check uses a different locale than my interactive session.  The issue resided in this: R interactively used:LC_COLLATE=en_US.UTF-8; R CMD check used: LC_COLLATE=C; https://stackoverflow.com/questions/42272119/r-cmd-check-fails-devtoolstest-works-fine
                        out.file = paste("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping"
                                         ,"/"
                                         ,species,"_"
                                         ,type.pair.name,".RData"
                                         ,sep=""
                                         )
                        print(out.file)
                        stop()
                        save(mapping.list,file=out.file)
               }
       }
}


  
```   
### species specific entrezid to ko
```{r}
load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/KO_ENTREZID.RData")
# mapping.table.all = mapping.table
# species = "hsa"
species = "mmu"
mapping.table = mapping.table.all[mapping.table.all[,"species"] == species,]
head(mapping.table)
tail(mapping.table)
out.file = paste("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping"
         ,"/"
         ,species,"_"
         ,"KO_ENTREZID.RData"
         ,sep=""
         )
print(out.file)
stop()
save(mapping.table,file=out.file)

```



## compounds

### pathwayCommons

extract sbgn compound id --> chebi from biopax:

script: 

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\pathway.commons\pathway.commons_SBGN.id.to.uniprot.chebi.using.biopax.R
input:

        pathwayCommons biopax files in folder: /media/sf_work/input/id.mapping/pathway.commons.10.biopax
output:

<!-- ###################################################################### -->
        
        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\pathway.commons\SBGNview.uniprot.chebi.to.SBGN.id.mapping.RData
<!-- ###################################################################### -->

#### lost reactome compounds 3/29/2019
```{r}


# library(xml2) 
biopax.file = "C:/xiaoxi/work/papers/SBGNview/package/data.package/pathway.commons/raw.data/v9/PathwayCommons9.reactome.BIOPAX.owl"
out.list = find.id.mapping.each.biopax.file.pathwayCommons10(biopax.file)
head(out.list$small.molecules.all)
dim(out.list$small.molecules.all)
cpds = out.list$small.molecules.all
length(unique(cpds$database.ids))
chebi.to.pc = cpds[,c("database.ids","id.q")]
colnames(chebi.to.pc) = c("chebi","pathwayCommons")
head(chebi.to.pc)
chebi.to.pc[,"chebi"] = gsub("CHEBI:","",chebi.to.pc[,"chebi"])
head(chebi.to.pc)

# merge added 
var = load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/chebi_pathwayCommons.RData")
var
chebi.to.pathwayCommons = mapping.list$compound$chebi_pathwayCommons
var = load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/pathwayCommons_pathway.id.RData")
var
pc.to.pathway = mapping.table
length(intersect(chebi.to.pc$pathwayCommons
                 ,chebi.to.pathwayCommons[,"pathwayCommons"]))
head(chebi.to.pc)
chebi.to.pc.m = as.matrix(chebi.to.pc)
head(chebi.to.pc.m)
chebi.to.pc.m = chebi.to.pc.m[,colnames(chebi.to.pathwayCommons)]
class(chebi.to.pc.m[1,1])
class(chebi.to.pc.m[1,2])
head(chebi.to.pc.m)
head(chebi.to.pathwayCommons)
dim(chebi.to.pc.m)
dim(chebi.to.pathwayCommons)
chebi.to.pc.all = rbind(chebi.to.pc.m,chebi.to.pathwayCommons)
dim(chebi.to.pc.all)
chebi.to.pc.all = unique(chebi.to.pc.all)
dim(chebi.to.pc.all)
dim(mapping.list$compound$chebi_pathwayCommons)
mapping.list$compound$chebi_pathwayCommons = chebi.to.pc.all
dim(mapping.list$compound$chebi_pathwayCommons)
save(mapping.list,file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/chebi_pathwayCommons.RData")

# genereate new chebi to pathway
head(pc.to.pathway)
head(chebi.to.pc.all)
pc.to.pathway.all = merge(pc.to.pathway
                          ,chebi.to.pc.all
                          ,all.y=T)
head(pc.to.pathway.all)
pc.to.pathway.all = pc.to.pathway.all[,c("chebi","pathway.id")]
dim(pc.to.pathway.all)
pc.to.pathway.all = unique(pc.to.pathway.all)
dim(pc.to.pathway.all)
head(pc.to.pathway.all)
var = load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/chebi_pathway.id.RData")
var
mapping.table = unique(mapping.table)
dim(mapping.table)
head(mapping.table)
mapping.table.without.reactome = mapping.table
mapping.table = rbind(mapping.table.without.reactome
                      ,pc.to.pathway.all)
head(mapping.table)
dim(mapping.table)
mapping.table = unique(mapping.table)
dim(mapping.table)
dim(mapping.table.without.reactome)
head(mapping.table)
save(mapping.table, file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/chebi_pathway.id.RData")



```
        
        
### metacyc

extract sbgn compound id --> chebi from biopax:

script:

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\metacyc/metacyc.SBGN.id.to.uniprot.uniref.chebi.R  ::  generate.biopax.id.to.chebi.or.metacyc()
        
result: 

<!-- ###################################################################### -->

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\metacyc/mapping.tables/compound.to.chebi.mapping.all.RData
<!-- ###################################################################### -->

### other compounds from unichem to pathwayCommons/metacyc compounds

1. Download unichem mapping tables to chebi and change names from "level..." to compound name:

script:

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\unichem\change.pairwise.mapping.file.name.pl
        
input: 
        
        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\unichem\unichem\ChEBI
        
output: 

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\unichem\unichem\ChEBI.db.names
        
        
2. merge tables:  metacyc/pathwayCommons compound id --> unichem ids

script:

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\all.ids.to.pathwayCommons.metacyc.R
        all.compounds.to.chebi.pathwayCommons.metacyc
        
result:

        
<!-- ###################################################################### -->

        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\pathway.commons\pathwayCommons.mappings.all.RData
        C:\xiaoxi\work\papers\SBGNview\package\id.mapping\id.mapping\metacyc\metacyc.SBGN.mappings.all.RData
<!-- ###################################################################### -->

### compound names

#### compound names to unichem IDs
script 

        C:\xiaoxi\work\papers\SBGNview\package\SBGN-ML.files\R\all.ids.to.pathwayCommons.metacyc.R :: all.compounds.to.compound.name
        
#### cpd name to metacyc id, pathwayCommons id
```{r}
chebi.to.pathwayCommons = paste(in.folder,"/chebi_pathwayCommons.RData",sep="")
fn = load(chebi.to.pathwayCommons )
head(mapping.list$compound$chebi_pathwayCommons)
chebi.to.pathwayCommons = mapping.list$compound$chebi_pathwayCommons
head(chebi.to.cpd.name)
head(chebi.to.pathwayCommons)
merged = merge(chebi.to.cpd.name,chebi.to.pathwayCommons,by="chebi",all=F)
head(merged)
mapping.table = merged[,c("CompoundName","pathwayCommons")]
dim(mapping.table)
mapping.table = unique(mapping.table)
dim(mapping.table)
head(mapping.table)
tail(mapping.table)
save(mapping.table,file = paste(in.folder,"/CompoundName_pathwayCommons.RData",sep=""))

# cpd name to metacyc id
chebi.to.metacyc = paste(in.folder,"/chebi_metacyc.SBGN.RData",sep="")
fn = load(chebi.to.metacyc )
head(mapping.list$compound$chebi_metacyc.SBGN)
chebi.to.metacyc = mapping.list$compound$chebi_metacyc.SBGN
head(chebi.to.cpd.name)
head(chebi.to.metacyc)
merged = merge(chebi.to.cpd.name,chebi.to.metacyc,by="chebi",all=F)
head(merged)
mapping.table = merged[,c("CompoundName","metacyc.SBGN")]
dim(mapping.table)
mapping.table = unique(mapping.table)
dim(mapping.table)
head(mapping.table)
tail(mapping.table)
save(mapping.table,file = paste(in.folder,"/CompoundName_metacyc.SBGN.RData",sep=""))



```


## From a list containing all mappings, Generate individual small mapping files
```{r}
generate.individual.file = function(input.list,gene.or.compound,output.folder){
        pairs = names(input.list)
        print(pairs)
        for (i in 1:length(pairs)){
                mapping.list = list()
                mapping.list[[gene.or.compound]] = list()
                pair = pairs[i]
                print(pair)
                mapping.list[[gene.or.compound]][[pair]] =  input.list[[pair]][!is.na(input.list[[pair]][,1]),]
                # head(ENTREZID_pathwayCommons)
                # tail(ENTREZID_pathwayCommons[[gene.or.compound]]$ENTREZID_pathwayCommons)
                # dim(ENTREZID_pathwayCommons[[gene.or.compound]]$ENTREZID_pathwayCommons)
                # print(table(ENTREZID_pathwayCommons[[gene.or.compound]]$ENTREZID_pathwayCommons[,3]))
                output.file = paste(output.folder,pair,".RData",sep="")
                print(output.file)
                save(mapping.list,file = output.file)
        }
}
output.folder = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/"

v.name = load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files//R/id.mapping/pathway.commons.and.metacyc/entrezid.to.metacyc.pathwayCommons.with.species.info.RData") # the metacyc table in this file only has species included by pathview, therefore it lost many other species
input.list = entrezid.to.metacyc.pathwayCommons.with.species.info
gene.or.compound = "gene"
generate.individual.file(input.list,gene.or.compound,output.folder)
# This file has all original species from reactome, mapped from UniProt2Reactome_PE_All_Levels.txt
load(file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/mapping.each.pair/for.use.all/both/pathwayCommons_UNIPROT.RData")
as.matrix(table(mapping.list$gene$pathwayCommons_UNIPROT[,"species"]))
head(mapping.list$gene$pathwayCommons_UNIPROT)
# check the original table
v.name = load("C:/xiaoxi/work/papers/SBGNview/package/id.mapping/id.mapping/pathway.commons/uniprot.to.pathwayCommons.all.species.list.RData")
as.matrix(table(uniprot.to.pathwayCommons.all.species.list$gene[,"species"]))

#################

v.name = load("C:/xiaoxi/work/papers/SBGNview/package/id.mapping/id.mapping/metacyc/metacyc.SBGN.mappings.all.RData")  # In this file, the metacyc genes are from only a subset of species included by pathview, therefore it lost many other species
length(unique((metacyc.SBGN.mappings.all$gene$metacyc.SBGN_UNIPROT[,1])))

#The full list of uniprot to metacyc should be:
load("C:/xiaoxi/work/papers/SBGNview/package/data.package/metacyc/merged.tables/uniprot.pathway.sbgnBiopaxProtein.RData")
length(unique(uniprot.pathway.sbgnBiopaxProtein[,1]))
head(uniprot.pathway.sbgnBiopaxProtein)
metacyc.SBGN_UNIPROT = uniprot.pathway.sbgnBiopaxProtein[,c("Uniprot","sbgnBiopaxProtein")]
colnames(metacyc.SBGN_UNIPROT) = c("UNIPROT","metacyc.SBGN")
metacyc.SBGN_UNIPROT$species = "No.data"
metacyc.SBGN_UNIPROT = metacyc.SBGN_UNIPROT[!is.na(metacyc.SBGN_UNIPROT[,1]),]
dim(metacyc.SBGN_UNIPROT[is.na(metacyc.SBGN_UNIPROT[,1]),])
dim(metacyc.SBGN_UNIPROT)
length(unique(metacyc.SBGN_UNIPROT[,1]))
mapping.list = list()
mapping.list$gene = list()
mapping.list$gene$metacyc.SBGN_UNIPROT = metacyc.SBGN_UNIPROT
save(mapping.list,file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/mapping.each.pair/for.use.all/both/metacyc.SBGN_UNIPROT.RData")
# try to get species information 
write.table(unique(mapping.list$gene$metacyc.SBGN_UNIPROT[,1]),"./metacyc.uniprot.txt",quote=F,row.names = F)

gene.or.compound = "gene"
input.list = metacyc.SBGN.mappings.all[[gene.or.compound]]
generate.individual.file(input.list,gene.or.compound,output.folder)

gene.or.compound = "compound"
input.list = metacyc.SBGN.mappings.all[[gene.or.compound]]
generate.individual.file(input.list,gene.or.compound,output.folder)

load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files//R/id.mapping/pathway.commons/pathwayCommons.mappings.all.RData")
gene.or.compound = "gene"
input.list = pathwayCommons.mappings.all[[gene.or.compound]]
generate.individual.file(input.list,gene.or.compound,output.folder)

gene.or.compound = "compound"
input.list = pathwayCommons.mappings.all[[gene.or.compound]]
generate.individual.file(input.list,gene.or.compound,output.folder)

```
Results are organized in folder :

        C:/xiaoxi\work\papers\SBGNview\package\SBGN-ML.files\data\mapping.each.pair

Find which ids can be mapped:

```{r}
mapped.ids = list()
mapped.ids$pathwayCommons.mappable.gene.ids = pathwayCommons.mappings.all$pathwayCommons.mappable.gene.ids
mapped.ids$pathwayCommons.mappable.compound.ids = pathwayCommons.mappings.all$pathwayCommons.mappable.compound.ids
mapped.ids$metacyc.mappable.gene.ids = metacyc.SBGN.mappings.all$metacyc.mappable.gene.ids
mapped.ids$metacyc.mappable.compound.ids = metacyc.SBGN.mappings.all$metacyc.mappable.compound.ids

mapped.ids$metacyc.mappable.gene.ids = c(mapped.ids$metacyc.mappable.gene.ids,"ENTREZID" )
mapped.ids$pathwayCommons.mappable.gene.ids = c(mapped.ids$pathwayCommons.mappable.gene.ids,"ENTREZID" )
mapped.ids
save(mapped.ids,file="../data/mapped.ids.RData ")
```
# Molecule ID to pathway IDs
## metacyc node id to pathway id
```{r 2019_1_10}
library(SBGNview)
data("pathways.info")
head(pathways.info)
if.metacyc = pathways.info$database == "MetaCyc"
sum(if.metacyc)
metacyc.ids = pathways.info$pathway.id[if.metacyc]
head(metacyc.ids)
result.list = sbgn.nodes(
                input.sbgn = metacyc.ids
                ,sbgn.folder = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/SBGN/MetaCyc" )

for (i in 1:length(result.list)){
        i=1
        all.nodes = result.list[[i]]$all.nodes
        all.nodes = all.nodes[all.nodes[,"class"] %in% c("macromolecule","simple chemical"),]
        head(all.nodes)
        # I didn't generate metacyc to pathway.id, now generate this:
        metacyc.pathways = strsplit(as.character(all.nodes[,"id"],":@:"))
        metacyc.pathways = do.call(rbind,metacyc.pathways)
        metacyc.pathways = cbind(as.character(input.to.metacyc[,"node.id"]),metacyc.pathways)
        metacyc.pathways = metacyc.pathways[,c(1,3)]
        colnames(metacyc.pathways) = c("metacyc.SBGN","pathway.id")
        head(metacyc.pathways)
        mapping.table = metacyc.pathways
        head(mapping.table)
        tail(mapping.table)
        save(mapping.table,file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/mapping.each.pair/for.use.all/both/metacyc.SBGN_pathway.id.RData")

}

```


## other id types to pathway id: metacyc and pathwayCommons

```{r}

in.folder = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/mapping.each.pair/for.use.all/both"
pathwayCommons.files = list.files(in.folder,pattern="pathwayCommons",full.names = T) # previously, we already mapped ids to metacyc pathway ids, here we only need to add mapping to pathwayCommons ids
length(pathwayCommons.files)
pathwayCommons.files.names = list.files(in.folder,pattern="pathwayCommons",full.names = F)
pathwayCommons.to.pathway.id = pathwayCommons.files[grepl("pathway.id",pathwayCommons.files.names)]
pathwayCommons.to.pathway.id
fn = load(pathwayCommons.to.pathway.id)
head(pathwayCommons_pathway.id)

pathwayCommons.files = pathwayCommons.files[!grepl("pathway.id",pathwayCommons.files.names)]
pathwayCommons.files.names = pathwayCommons.files.names[!grepl("pathway.id",pathwayCommons.files.names)]
length(pathwayCommons.files)
length(pathwayCommons.files.names)
head(pathwayCommons.files.names)
pathwayCommons.files.names

# for( i in 1:length(pathwayCommons.files)){
for( i in c(26,47)){
        i=26
        i=47
        print(pathwayCommons.files.names[i])
        # i=7  # previously, only chebi didn't have mapping to pathway ids of pathwayCommons, so we only need to add chebi.to.pathwayCommons.pathway.id
        # find id type and parse metacyc, generate metacyc to pathway.id mapping
        load(pathwayCommons.files[i])
        id.to.pathwayCommons = mapping.list[[1]][[1]]
        id.type = setdiff(colnames(id.to.pathwayCommons),c("species","pathwayCommons"))
        
        type.pair.name = paste(sort(c(id.type,"metacyc.SBGN")),collapse="_")
        id.to.metacyc.file = paste(in.folder,"/",type.pair.name,".RData",sep="")
        # print(id.to.pathway.file)
        load(id.to.metacyc.file)
        print(id.type)
        idType.to.pathway.id = mapping.list[[1]][[1]]
        head(idType.to.pathway.id)
        tail(idType.to.pathway.id)
        dim(idType.to.pathway.id)
        idType.to.pathway.id = idType.to.pathway.id[!grepl("NA:@:NA",idType.to.pathway.id[,"metacyc.SBGN"]),]
        dim(idType.to.pathway.id)
        pathway.ids = do.call(rbind,strsplit(idType.to.pathway.id$metacyc.SBGN,":@:"))[,2]
        head(pathway.ids)
        length(pathway.ids)
        head(as.character(as.vector(idType.to.pathway.id[,id.type])))
        head(idType.to.pathway.id[,id.type])
        idType.to.pathway.id = cbind(as.character(as.vector(idType.to.pathway.id[,id.type])),pathway.ids)
        colnames(idType.to.pathway.id) =c(id.type,"pathway.id")
        head(idType.to.pathway.id)
        tail(idType.to.pathway.id)
        
        
        # parse pathwaycOMMONS, GENERATE IDTYPE TO PATHWAY ID MAPPING FOR PATHWAYcOMMONS
        head(pathwayCommons_pathway.id)
        head(id.to.pathwayCommons)
        merged = merge(pathwayCommons_pathway.id,id.to.pathwayCommons,by="pathwayCommons",all=F)
        head(merged)
        dim(merged)
        id.to.pathway = merged[,c(id.type,"pathway.id")]
        head(id.to.pathway)
        mapping.table = idType.to.pathway.id
        head(mapping.table)
        tail(mapping.table)
        mapping.table = rbind(mapping.table,id.to.pathway)
        head(mapping.table)
        tail(mapping.table)
        dim(mapping.table)
        id.to.pathway.file = paste(in.folder,"/",id.type,"_pathway.id.RData",sep="")
        id.to.pathway.file
        save(mapping.table,file = id.to.pathway.file)
        
        load(id.to.pathway.file)
        head(mapping.table)
        tail(mapping.table)
}

```
### compound name to pathway ids
```{r}
        
chebi.to.cpd.name.file = paste(in.folder,"/chebi_CompoundName.RData",sep="")
fn = load(chebi.to.cpd.name.file)
head(mapping.list$compound$chebi_CompoundName)
chebi.to.cpd.name = mapping.list$compound$chebi_CompoundName
head(chebi.to.cpd.name)
chebi.to.cpd.name[chebi.to.cpd.name[,1] == "64489",]
head(mapping.table)
chebi.to.pathway = mapping.table
head(chebi.to.cpd.name)
head(chebi.to.pathway)
merged = merge(chebi.to.cpd.name,chebi.to.pathway,by="chebi",all=F)
head(merged)
mapping.table = merged[,c("CompoundName","pathway.id")]
dim(mapping.table)
mapping.table = unique(mapping.table)
dim(mapping.table)
head(mapping.table)
tail(mapping.table)
save(mapping.table,file = paste(in.folder,"/CompoundName_pathway.id.RData",sep=""))
```

## metacrop
```{r}
library(SBGNview)
data("pathways.info")
    source("C:/xiaoxi/work/papers/SBGNview/package/SBGNview/R/parsing.utilities.R")
    source("C:/xiaoxi/work/papers/SBGNview/package/SBGNview/R/plot.utilities.R")
    source("C:/xiaoxi/work/papers/SBGNview/package/SBGNview/R/SBGN.to.SVG.R")
is.metacrop = pathways.info[,"database"]== "MetaCrop"
metacrop.ids = pathways.info[is.metacrop ,"pathway.id"]
node.list = sbgn.nodes(input.sbgn = c( metacrop.ids
                                       # ,metacyc.ids[1:2]
                                        )
                               )
node.list$`Alanine Degradation`$all.nodes
```
### compound name to pathway id

```{r}
load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/CompoundName_pathway.id.without.metacrop.RData")
dim(mapping.table)
head(mapping.table)
tail(mapping.table)
class(mapping.table[,1])
mapping.table[,1] = as.character(mapping.table[,1])
mapping.table[,2] = as.character(mapping.table[,2])

for (i in 1:length(node.list)){
        pathway = names(node.list)[i]
        label.to.class = node.list[[i]]$all.nodes[,c("class","label")]
        label.to.class = as.data.frame(label.to.class,stringsAsFactors=F)
        class(label.to.class)
        cpd.name.to.pathway = cbind(subset(label.to.class,class=="simple chemical",select="label"),pathway)
        cpd.name.to.pathway
        colnames(cpd.name.to.pathway) = c("CompoundName","pathway.id")
        mapping.table = rbind(mapping.table,cpd.name.to.pathway)
}
dim(mapping.table)
mapping.table = unique(mapping.table)
dim(mapping.table)
tail(mapping.table)
head(mapping.table)
class(mapping.table[,1])
class(mapping.table[,2])
save(mapping.table,file="C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/CompoundName_pathway.id.RData")
```

### EC number to pathway id

```{r}
load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/ENZYME_pathway.id.without.METACROP.RData")
dim(mapping.table)

for (i in 1:length(node.list)){
        pathway = names(node.list)[i]
        label.to.class = node.list[[i]]$all.nodes[,c("class","label")]
        label.to.class = as.data.frame(label.to.class,stringsAsFactors=F)
        class(label.to.class)
        ec.to.pathway = cbind(subset(label.to.class,class=="macromolecule",select="label"),pathway)
        ec.to.pathway
        colnames(ec.to.pathway) = c("ENZYME","pathway.id")
        mapping.table = rbind(mapping.table,ec.to.pathway)
}
mapping.table = unique(mapping.table)
dim(mapping.table)
tail(mapping.table)
head(mapping.table)
class(mapping.table[,1])
save(mapping.table,file="C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/ENZYME_pathway.id.RData")

# clean gene names
m.file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/ENZYME_pathwayCommons.RData"
clean.ez.names = function(m.file){
        aa = load(file=m.file)
        if(is.list(get(aa))){
                mapping.table = get(aa)[[1]][[1]]
        }
        head(mapping.table)
        ezs = mapping.table[,1]
        head(ezs)
        if.name = grepl(pattern = "[a-zA-Z]",ezs)
        head(if.name)
        sum(if.name)
        length(ezs)
        dim(mapping.table)
        mapping.table = mapping.table[!if.name,]
        dim(mapping.table)
        save(mapping.table,file=m.file)
}

clean.ez.names("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/ENZYME_pathway.id.RData")
        
```


#number of kos mapped in each species-specific pathway
```{r,number of kos mapped in each species-specific pathway}


use.uniprot.species.info = function(){ # this method is not good. because some KOs have genes in a species but don't have uniprot in the species, thus lose information. examples: ko:K00002	ptr:741418;  http://rest.kegg.jp/conv/uniprot/ptr:741418+mmu:58810
        # number of kos mapped in species specific pathway
        load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/mapping.each.pair/for.use.all/both/pathwayCommons_pathway.id.RData")
        pathwayCommons_pathway.id.protein = pathwayCommons_pathway.id[grepl("Protein_",pathwayCommons_pathway.id[,"pathwayCommons"]),]
        dim(pathwayCommons_pathway.id.protein)
        pathwayCommons_pathway.id.protein = pathwayCommons_pathway.id.protein[!grepl("info",pathwayCommons_pathway.id.protein[,"pathwayCommons"]),]
        dim(pathwayCommons_pathway.id.protein)
        pathwayCommons_pathway.id.protein = pathwayCommons_pathway.id.protein[!grepl("Complex",pathwayCommons_pathway.id.protein[,"pathwayCommons"]),]
        dim(pathwayCommons_pathway.id.protein)
        pathwayCommons_pathway.id.protein = pathwayCommons_pathway.id.protein[!grepl("state",pathwayCommons_pathway.id.protein[,"pathwayCommons"]),]
        dim(pathwayCommons_pathway.id.protein)
        head(pathwayCommons_pathway.id.protein)
        pathway.protein.count = table(pathwayCommons_pathway.id.protein[,"pathway.id"])
        head(pathway.protein.count)
        load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/mapping.each.pair/for.use.all/both/pathwayCommons_UNIPROT.RData")
        head(mapping.list$gene$pathwayCommons_UNIPROT)
        uniprot.to.species = unique(mapping.list$gene$pathwayCommons_UNIPROT[,c("pathwayCommons","species")])
        head(uniprot.to.species)
        species.to.pathway = merge(uniprot.to.species,pathwayCommons_pathway.id.protein,by="pathwayCommons",all=F)
        head(species.to.pathway)
}

ko.to.gene = read.table("C:/xiaoxi/work/input/id.mapping/kegg.related/kegg.ortholog.100.download.slice/kegg.ko.to.all.species.entrez.tsv",header=F,stringsAsFactors = F,quote="",sep="\t")
head(ko.to.gene)
ko.to.gene[,1] = gsub("ko:","",ko.to.gene[,1])
species = strsplit(ko.to.gene[,2],":")
species = do.call(rbind,species)
head(species)
species = species[,1]
ko.to.species = data.frame(KO=ko.to.gene[,1],species = species,stringsAsFactors = F)
dim(ko.to.species)
ko.to.species = unique(ko.to.species)
dim(ko.to.species)
head(ko.to.species)


load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/id.mapping/mapping.each.pair/for.use.all/both/KO_pathway.id.RData")
ko.to.pathway.id = mapping.table
head(ko.to.pathway.id)
dim(ko.to.pathway.id)
ko.to.pathway.id = unique(ko.to.pathway.id)
dim(ko.to.pathway.id)
species.to.pathway = merge(ko.to.species,ko.to.pathway.id,by="KO",all=F)
dim(species.to.pathway)
species.to.pathway = unique(species.to.pathway)
head(species.to.pathway)
pathway.protein.count = table(ko.to.pathway.id[,"pathway.id"])
head(pathway.protein.count)
length(pathway.protein.count)

# species.pathway.id = aggregate(
#         species.to.pathway[,"pathwayCommons"]
#         ,list(species = species.to.pathway[,"species"]
#               ,pathway = species.to.pathway[,"pathway.id"]
#               )
#         ,length
# )
species.pathway = paste(species.to.pathway[,"pathway.id"],species.to.pathway[,"species"],sep="@")
head(species.pathway)
species.pathway.count = table(species.pathway)
head(species.pathway.count)
length(species.pathway.count)
species.pathway.count.n = data.frame(pathway = as.character(names(species.pathway.count)),count = species.pathway.count,stringsAsFactors = F)
head(species.pathway.count.n[!is.character(species.pathway.count.n$pathway),])
head(class(as.character(names(species.pathway.count))))
head(species.pathway.count.n$pathway)
head(class(species.pathway.count.n$pathway))
species.pathway.name = do.call(rbind,strsplit(species.pathway.count.n$pathway,"@"))
head(species.pathway.name)

pathway.protein.count.v = pathway.protein.count[species.pathway.name[,1]]
head(pathway.protein.count.v)
species.pathway.name = as.data.frame(species.pathway.name)

species.pathway.id = species.pathway.name
species.pathway.id$ref.pathway.protein.count = pathway.protein.count.v
head(species.pathway.id)
colnames(species.pathway.id) = c("pathway","species","ref.pathway.protein.count")
pct.mapped.species.pathway = as.numeric(species.pathway.count) / pathway.protein.count.v
head(pct.mapped.species.pathway)
species.pathway.id$pct.mapped.species.pathway = pct.mapped.species.pathway
species.pathway.id$species.pathway.protein.count = species.pathway.count
species.pathway.id$ref.pathway.protein.count = pathway.protein.count.v
head(species.pathway.id)
head(species.pathway.id[species.pathway.id[,"species"] == "hsa",])
min(species.pathway.id[species.pathway.id[,"species"] == "hsa","pct.mapped.species.pathway"])
hist(species.pathway.id[species.pathway.id[,"species"] == "hsa","pct.mapped.species.pathway"])
hist(species.pathway.id[species.pathway.id[,"species"] == "mmu","pct.mapped.species.pathway"])
hist(species.pathway.id[species.pathway.id[,"species"] == "ptr","pct.mapped.species.pathway"])
range(species.pathway.id[species.pathway.id[,"species"] == "mmu","pct.mapped.species.pathway"])
range(species.pathway.id[species.pathway.id[,"species"] == "ptr","pct.mapped.species.pathway"])
mean(species.pathway.id[species.pathway.id[,"species"] == "mmu","pct.mapped.species.pathway"])
mean(species.pathway.id[species.pathway.id[,"species"] == "ptr","pct.mapped.species.pathway"])
median(species.pathway.id[species.pathway.id[,"species"] == "mmu","pct.mapped.species.pathway"])
median(species.pathway.id[species.pathway.id[,"species"] == "ptr","pct.mapped.species.pathway"])
length(species.pathway.id[species.pathway.id[,"species"] == "mmu","pct.mapped.species.pathway"])
length(species.pathway.id[species.pathway.id[,"species"] == "ptr","pct.mapped.species.pathway"])
head(species.pathway.id)
max(species.pathway.id$pct.mapped.species.pathway)
hist(species.pathway.id$pct.mapped.species.pathway)
pathway.species.pct_Mapped = species.pathway.id
save(pathway.species.pct_Mapped,file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/pathway.species.pct_Mapped.RData")

load("C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/pathway.species.pct_Mapped.RData")
head(pathway.species.pct_Mapped)
pathway.species.pct_Mapped[pathway.species.pct_Mapped[,"species"] == "hsa","pct.mapped.species.pathway"] = 1 # not all pathway nodes(pathwayCommons IDs) have uniprot, thus species information. This is why some human pathways don't have 100% completeness. Here we change them to 1 so cutoff filtering can work properly.
min(pathway.species.pct_Mapped[pathway.species.pct_Mapped[,"species"] == "hsa","pct.mapped.species.pathway"])
head(pathway.species.pct_Mapped)
hist(pathway.species.pct_Mapped[pathway.species.pct_Mapped[,"species"] == "mmu","pct.mapped.species.pathway"])
hist(pathway.species.pct_Mapped[pathway.species.pct_Mapped[,"species"] == "hsa","pct.mapped.species.pathway"])
range(pathway.species.pct_Mapped[pathway.species.pct_Mapped[,"species"] == "hsa","pct.mapped.species.pathway"])
save(pathway.species.pct_Mapped,file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/pathway.species.pct_Mapped.RData")


```
### pathway specific cutoff for completeness
```{r}
load(file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/pathway.species.pct_Mapped.RData")
head(pathway.species.pct_Mapped)
# table(pathway.species.pct_Mapped[1:1000,"pathway"] )
all.species = as.character(unique(pathway.species.pct_Mapped$species))
length(all.species)
head(all.species)

# library(plyr)
cutoff.to.fs = function(pathway.species.pct_Mapped.in){
# pathway.species.pct_Mapped.in = pathway.species.pct_Mapped
        pathway.species.pct_Mapped.in = as.data.frame(pathway.species.pct_Mapped.in,stringsAsDataframe= F)
        head(pathway.species.pct_Mapped)
        # colnames(pathway.species.pct_Mapped.in)[1] = "species"
        out = matrix(ncol = 4,nrow = 0)
        print(dim(pathway.species.pct_Mapped.in))
        pathway.cutoff.anova = by(
        # pathway.cutoff.anova = ddply(
                pathway.species.pct_Mapped.in
                ,as.factor(pathway.species.pct_Mapped.in$pathway)
                ,function(df){
                        covs = df$pct.mapped.species.pathway
                        pathway.species = as.character(unique(df$species))
                        head(pathway.species)
                        length(pathway.species)
                        head(covs)
                        cutoffs = unique(covs)
                        mean.covs = mean(covs)
                        length(cutoffs)
                        cutoffs = cutoffs[order(cutoffs)]
                        if(length(cutoffs)<5){
                                cutoff.anova = c(pathway = as.character(df$pathway[1])
                                                 ,cutoff = mean.covs
                                                 ,stat = paste("number of uniq coverage values:",length(unique(covs)),".  Not enough for anova analysis, using average coverage as cutoff",sep="")
                                                 ,P=paste(c("uniq coverage values:",cutoffs),collapse = "\t")
                                                 )
                                out<<- rbind(out,t(cutoff.anova))
                                return()
                        }
                        cutoffs = cutoffs[3:(length(cutoffs)-2)]
                        head(cutoffs)
                        
                        cutoff.anova = sapply(cutoffs
                               ,function(cutoff){
                                        if.pathway.exist = rep("no",times=length(covs))
                                        if.pathway.exist[covs>=cutoff] = "yes"
                                        cutoff.assign = data.frame(cov = as.vector(covs), if.exist = if.pathway.exist,stringsAsFactors = F)
                                        # cutoff.assign$cov = cutoff.assign$cov + runif(nrow(cutoff.assign),0,1)/1000000
                                        cutoff.assign
                                        tst = try(oneway.test(covs~if.exist,data=cutoff.assign,var.equal = T))
                                        if(class(tst) == "try-error"){
                                                print(head(df))
                                                print(head(cutoff.assign))
                                                print(dim(df))
                                                print(cutoffs)
                                                print(cutoff)
                                                print(dim(cutoff.assign))
                                                boxplot(cutoff.assign$cov ~ cutoff.assign$if.exist)
                                                print(table(cutoff.assign$if.exist))
                                                stop()
                                        }
                                        f.p = c(pathway = as.character(df$pathway[1]),cutoff = cutoff, stat = tst$statistic,P =tst$p.value)
                                        f.p
                                        return(f.p)
                               }
                               )
                        out<<- rbind(out,t(cutoff.anova))
                }
        )
        print(head(out))
        dim(out)
        return(out)
}

out = cutoff.to.fs(pathway.species.pct_Mapped)

```



Some pathways have <=4 unique coverage values across all species, we use the average values of coverage in all species as cutoff
```{r}

head(out)
colnames(out)[3] = "stat.F"
out.mean.as.cutoff= out[grepl("number",out[,"stat.F"]),]
head(out.mean.as.cutoff)
dim(out.mean.as.cutoff)
```
For pathways with enough unique coverages, we try every coverage value as cutoff to assign "exist" and "not exist" for all species-specific pathways, then use anova to calculate F statistics and P value for the difference between "exist" and "not exist" group under this cutoff. Then we select the cutoff that generates largest F value between the two groups.
```{r}
out.anova.as.cutoff = out[!grepl("number",out[,"stat.F"]),] 
head(out.anova.as.cutoff)
dim(out.anova.as.cutoff)
cutoffs.to.anova.f = out.anova.as.cutoff
save(cutoffs.to.anova.f,file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/cutoffs.to.anova.f.RData")

out.anova.max.F.as.cutoff = matrix(ncol=4,nrow=0)
if.max.F = by(
        as.data.frame(out.anova.as.cutoff,stringsAsFactors = F)
        ,as.factor(out.anova.as.cutoff[,"pathway"])
        ,function(df){
                f.stats = as.numeric(df[,"stat.F"])
                # print(f.stats)
                if.max = f.stats == max(f.stats)
                out.anova.max.F.as.cutoff <<- rbind(out.anova.max.F.as.cutoff
                                                    ,as.vector(df[if.max,])
                                                    )
                
        }
)

out.anova.max.F.as.cutoff[,2] = as.numeric(out.anova.max.F.as.cutoff[,2])
out.anova.max.F.as.cutoff[,3] = as.numeric(out.anova.max.F.as.cutoff[,3])
out.anova.max.F.as.cutoff[,4] = as.numeric(out.anova.max.F.as.cutoff[,4])
head(out.anova.max.F.as.cutoff ) 
dim(out.anova.max.F.as.cutoff ) 
plot(density(out.anova.max.F.as.cutoff$cutoff),main = "Distribution of species-specific pathway completeness cutoffs")
```
Check if all cutoffs generate grouping that showed significant difference between the  "exist" and "not exist"  groups
```{r}
max(out.anova.max.F.as.cutoff$P)
```
Save all cutoff information
```{r}
all.cutoff.info = rbind(out.anova.max.F.as.cutoff,out.mean.as.cutoff)
all.cutoff.info = as.data.frame(all.cutoff.info)
all.cutoff.info[,"cutoff"] = as.numeric( all.cutoff.info[,"cutoff"])
colnames(all.cutoff.info)[3:4] = c("ANOVA.F.stat","ANOVA.P")
head(all.cutoff.info)
pathway.completeness.cutoff.info = all.cutoff.info
plot(density(all.cutoff.info$cutoff),main = "Distribution of species-specific pathway completeness cutoffs.all")
save(pathway.completeness.cutoff.info,file = "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/pathway.completeness.cutoff.info.RData")
load( "C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/pathway.completeness.cutoff.info.RData")
```

Some pathways have many(sometimes all) species-specific pathways with 100% completeness, take a look at them.
```{r}
large.cutoff.value = all.cutoff.info[all.cutoff.info$cutoff > 0.9,]
head(large.cutoff.value)
dim(large.cutoff.value)
species.info = pathway.species.pct_Mapped[pathway.species.pct_Mapped[,"pathway"] %in% large.cutoff.value[,"pathway"],]
head(species.info)
n.species = table(as.character(species.info[,"pathway"]))
head(n.species)
plot(density(n.species),main = "Number of species where pathways with large cutoff are detected")
plot(hist(n.species),main = "Number of species where pathways with large cutoff are detected")
plot(hist(n.species[n.species>1],breaks = seq(1,max(n.species),by=5)),main = "Number of species where pathways with large cutoff are detected")

# library(ggplot2)
# p = qplot(n.species[n.species>1]
p = qplot(n.species
      ,geom="histogram"
      ,binwidth = 1
      ,main = "Number of species where pathways with large cutoff are detected(completeness >0)\n Pathways with large cutoff(>0.9): they have large average completeness and small variation of completeness across species"
      ,ylab = "count"
      )+ scale_y_log10() #+ scale_x_log10()
print(p)

table(as.character(n.species))
sum(table(as.character(n.species)))
head(n.species)
max(n.species)
n.species[pathway.largest.species.n]

sum(n.species>0)
pathway.largest.species.n = names(n.species)[n.species == max(n.species)]
pathway.largest.species.n
head(species.info[species.info[,"pathway"] == pathway.largest.species.n,])
dim(species.info[species.info[,"pathway"] == pathway.largest.species.n,])
range(species.info[species.info[,"pathway"] == pathway.largest.species.n,"pct.mapped.species.pathway"])
length(unique((species.info[species.info[,"pathway"] == pathway.largest.species.n,"species"])))
head(unique((as.character(species.info[species.info[,"pathway"] == pathway.largest.species.n,"species"]))))
length(unique((as.character(species.info[species.info[,"pathway"] == pathway.largest.species.n,"species"]))))
```

## for each species specific pathway, how many proteins does it have
```{r}
ko.to.pathway.file = "C:/xiaoxi/work/input/id.mapping/kegg/ko.to.pathway.tsv"
download.file("http://rest.kegg.jp/link/ko/pathway",destfile = ko.to.pathway.file )
all.ko.to.pathway = read.table(ko.to.pathway.file,header=F,stringsAsFactors = F,sep="\t")
head(all.ko.to.pathway)
pathway.ko.count = table(all.ko.to.pathway[,1])
head(pathway.ko.count)
tail(pathway.ko.count)


all.species = read.table("C:/xiaoxi/work/input/id.mapping/kegg/kegg.ko.to.all.species.tsv",stringsAsFactors = F,header=F)
head(all.species)
all.species = unique(all.species[,1])
length(all.species)

out.all = matrix(ncol=4,nrow=0)
out.all
head(out.all)
dim(out.all)

for(i in 1:length(all.species)){
        # i = 1
        cat("\n\n",i)
        print(all.species[i])
        dlink = paste("http://rest.kegg.jp/link/",all.species[i],"/pathway",sep="")
        loc.file = paste("C:/xiaoxi/work/input/id.mapping/kegg/kegg.species.to.gene/",all.species[i],".tsv")
        cov.file = paste(loc.file,".pct.cover.tsv")
        if(file.exists(cov.file)){
                next()
        }
        if.downloaded = try(download.file(url = dlink,destfile = loc.file))
        if(class(if.downloaded) == "try-error"){
                next()
        }
        pathway.to.gene = read.table(loc.file,header=F,sep="\t",stringsAsFactors = F)
        ct = table(pathway.to.gene[,1])
        ct[ct==min(ct)]
        ko.count.pathway = pathway.ko.count[gsub(all.species[i],"map",names(ct))]
        percent.species.gene.vs.ko.in.pathway = ct/ko.count.pathway
        out = data.frame(
                pathway = names(ct)
                ,n.gene.species.pathway = as.vector(ct)
                ,n.ko.reference.pathway = as.vector(ko.count.pathway)
                ,percent = as.vector(percent.species.gene.vs.ko.in.pathway)
        )
        out = out[order(out[,"percent"]),]
        out.all = rbind(out.all,out)
        write.table(out,cov.file,sep="\t",row.names=F)
}

write.table(out.all,"C:/xiaoxi/work/input/id.mapping/kegg/speciesPathwayMappedGene.refKo.numbers.tsv",sep="\t",row.names=F)

head(out.all)
species.info = gsub("[0-9]|path:","",as.character(out.all[,1]))
head(species.info)
head(out.all)
dim(out.all)

pathway.info = gsub("path:[a-z]+","map",as.character(out.all[,1]))
head(pathway.info)
out.all.backup = out.all
out.all = out.all[!is.na(out.all[,"percent"]),]
dim(out.all[is.na(out.all[,"percent"]),])

max.cov.pathway = by(out.all,as.factor(pathway.info),
                      function(df){
                              max.percent.df = df[df[,"percent"] == max(df[,"percent"]),]
                              max.percent.df = max.percent.df[1,]
                      }
              )
max.cov.pathway[[1]]

max.cov.pathway = do.call(rbind,max.cov.pathway )
head(max.cov.pathway)
                     

min.gene = by(out.all,as.factor(species.info),
              function(df){
                      df[df[,"n.gene.species.pathway"] == min(df[,"n.gene.species.pathway"]),]
              }
              )
min.gene = do.call(rbind,min.gene)
min.gene = min.gene[order(min.gene[,"percent"]),]
head(min.gene)
range(min.gene[,2])
write.table(min.gene,"C:/xiaoxi/work/input/id.mapping/kegg/pathways.with.minimum.cov.tsv",sep="\t",row.names=F)

head(out.all[is.na(out.all[,4]),])

min.cov = by(out.all,as.factor(species.info),
              function(df){
                      df[ df[,"percent"] == min(df[,"percent"]),]
              }
              )
min.cov = do.call(rbind,min.cov)
head(min.cov)
min.cov[,4]
range(min.cov[,2])
```



### predict pathway using ape::ace
```{r,ace}
library(ape)
data(bird.orders)
x <- rnorm(23)
class(bird.orders)
bird.orders
bird.orders$tip.label
str(bird.orders)
ace(x, bird.orders)


# how to map kegg genome code to species tree:
# from https://www.genome.jp/kegg/catalog/org_list4.html, get all taxa ids. result: "kegg.all.phyliptree.phy"
# use taxa ids to generate tree: https://www.ncbi.nlm.nih.gov/Taxonomy/CommonTree/wwwcmt.cgi
# ape's read.tree won't recogonize the taxa nmes in the file downloaded directly, because the names were quoted with'
# change the following characters in "kegg.all.phyliptree.phy" before use "read.tree", otherwise spece will be removed by "ape": remove ' ()in taxa names; change space " " to "_". then read the tree. result: a tree object with modified name: all.kegg.tre

# map the taxa names in the tree file to taxa id using: https://www.ncbi.nlm.nih.gov/Taxonomy/TaxIdentifier/tax_identifier.cgi . some taxa names won't be recogonized by this website, so need to handle manually: result: names.all.phyliptree.to.id.txt
# manually modify names.all.phyliptree.to.id.txt, so it will be same as the modified names in the tree object all.kegg.tre. result: ncbi.name.to.id with modified name
# use ncbi.name.to.id to map tip labels in all.kegg.tre to ncbi id
# use kegg webpage to map kegg code to ncbi id: https://www.genome.jp/kegg/tool/map_taxonomy.html: result:



# change the following characters in "kegg.all.phyliptree.phy" before use "read.tree", otherwise spece will be removed by "ape": remove ' ()in taxa names; change space " " to "_". then read the tree. result: a tree object with modified name: all.kegg.tre
library(ape)
all.kegg.tre = read.tree(file="C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/trees/kegg.all.phyliptree.phy")
str(all.kegg.tre)
head(all.kegg.tre$edge)
head(all.kegg.tre$tip.label)
tail(all.kegg.tre$tip.label)
tail(all.kegg.tre$node.label)
table(all.kegg.tre$tip.label)
tip.labels = all.kegg.tre$tip.label

# input.kegg =  read.table(file="C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/trees/kegg.taxa.names.to.ids.txt",sep="\t",stringsAsFactors = F,quote="",header=F)
# head(input.kegg)
# input.kegg.taxa.names.tr = gsub(" ","_",input.kegg[,1])
# input.kegg.taxa.names.tr = gsub("\\(.+\\)","",input.kegg.taxa.names.tr,perl=T)
# input.kegg.taxa.names.tr = gsub("_+","_",input.kegg.taxa.names.tr,perl=T)
# input.kegg.taxa.names.tr = gsub("\\[","",input.kegg.taxa.names.tr,perl=T)
# input.kegg.taxa.names.tr = gsub("\'","",input.kegg.taxa.names.tr,perl=T)
# input.kegg.taxa.names.tr = gsub("\\]","",input.kegg.taxa.names.tr,perl=T)
# input.kegg.taxa.names.tr[grepl("haematococca",input.kegg.taxa.names.tr,ignore.case = T)]
# input.kegg[grepl("oleophilus",input.kegg[,1],ignore.case = T),]
# input.kegg.taxa.names.tr[grepl("oleophilus",input.kegg.taxa.names.tr,ignore.case = T)]
# tip.labels[grepl("oleophilus",tip.labels,ignore.case = T)]
# length(intersect(input.kegg.taxa.names.tr,all.kegg.tre$tip.label))
# length(input.kegg.taxa.names.tr)
# head(setdiff(all.kegg.tre$tip.label,input.kegg.taxa.names.tr))

# map the taxa names in the tree file to taxa id using: https://www.ncbi.nlm.nih.gov/Taxonomy/TaxIdentifier/tax_identifier.cgi . some taxa names won't be recogonized by this website, so need to handle manually: result: names.all.phyliptree.to.id.txt
## extract names
fileName <- 'C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/trees/kegg.all.phyliptree.phy'
all.kegg.tre.char = readChar(fileName, file.info(fileName)$size)
head(all.kegg.tre.char)
length(all.kegg.tre.char)
all.kegg.tre.char[1]
all.kegg.tre.char = strsplit(all.kegg.tre.char,"'")[[1]]
length(all.kegg.tre.char)
head(all.kegg.tre.char)
tail(all.kegg.tre.char)
t.names = all.kegg.tre.char[seq(2,(length(all.kegg.tre.char)-1),by=2)]
head(t.names)
tail(t.names)
write.table(t.names,"C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/trees/names.all.phyliptree.phy",col.names = F,row.names = F,quote=F,sep="\t")
# use ncbi to map taxa names to ids
ncbi.name.to.id =  read.table(file="C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/trees/names.all.phyliptree.to.id.txt",sep="\t",stringsAsFactors = F,quote="",header=T)
head(ncbi.name.to.id)
length(setdiff(tip.labels,ncbi.name.to.id[,1]))
head(setdiff(tip.labels,ncbi.name.to.id[,1]))
setdiff(tip.labels,ncbi.name.to.id[,1])

library(stringr)
# kegg.three.letters code to taxa id: use: https://www.genome.jp/kegg/tool/map_taxonomy.html.  Download htext. RESULT: 
kegg.code.to.taxa =  scan(file="C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/trees/kegg.code.to.ncbi.id.keg",sep="\n",quote="",comment.char = "",what="character")
head(kegg.code.to.taxa)

kegg.code.to.taxa.out = matrix(ncol=2,nrow=0)
for(i in 1:length(kegg.code.to.taxa)){
        if(grepl("TAX:",kegg.code.to.taxa[i])){
                taxa.string = kegg.code.to.taxa[i]
                tax.id = str_extract(taxa.string,"\\[TAX:(.+)\\]")
                tax.id = gsub("\\[TAX:","",tax.id)
                tax.id = gsub("\\]","",tax.id)
                tax.id
                tax.rank = str_split(taxa.string,"\\s+",n=2)[[1]][1]
                tax.rank
                tax.rank.children = toupper(letters)[which(toupper(letters) == tax.rank)+1]
                tax.rank.children
                for(j in (i+1):length(kegg.code.to.taxa)){
                        child.string = kegg.code.to.taxa[j]
                        child.rank = str_split(child.string,"\\s+",n=2)[[1]][1]
                        if(child.rank == tax.rank.children){
                                child.code = str_split(child.string,"\\s+",n=3)[[1]][2]
                                child.code
                                tax.id
                                kegg.code.to.taxa.out = rbind(kegg.code.to.taxa.out,c(child.code,tax.id))
                        }else{
                                break()
                        }
                }
        }
}
head(kegg.code.to.taxa.out)
head(ncbi.name.to.id)
head(all.kegg.tre$tip.label)
class(kegg.code.to.taxa.out[,1])
class(kegg.code.to.taxa.out[,2])
class(ncbi.name.to.id[,2])
dim(kegg.code.to.taxa.out)
colnames(kegg.code.to.taxa.out) = c("kegg.code","taxa.id")
write.table(kegg.code.to.taxa.out,"C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/trees/kegg.code.to.taxa",row.names = F,quote=F,sep="\t")

head(ncbi.name.to.id)
ncbi.name.to.id$taxid = as.character(ncbi.name.to.id$taxid)
kegg.code.to.ncbi.name = merge(kegg.code.to.taxa.out,ncbi.name.to.id,by.x="taxa.id",by.y = "taxid",all.y=T)
dim(kegg.code.to.ncbi.name)
# kegg.code.to.ncbi.name = kegg.code.to.ncbi.name[!duplicated(kegg.code.to.ncbi.name$name),]
dim(kegg.code.to.ncbi.name)
head(kegg.code.to.ncbi.name)
row.names(kegg.code.to.ncbi.name) = kegg.code.to.ncbi.name[,"name"]
str(all.kegg.tre)
head(all.kegg.tre$edge)
head(all.kegg.tre$tip.label)
tail(all.kegg.tre$tip.label)
tail(all.kegg.tre$node.label)
table(all.kegg.tre$tip.label)
length(setdiff(tip.labels,kegg.code.to.ncbi.name$name))
length(setdiff(kegg.code.to.ncbi.name$name,tip.labels))




# # load saved kegg taxa name to id, note: this name is species name, which could be different from the name in the following "3letters to name" mapping(which is genome name)
# kegg.name.to.taxa.id =  read.table(file="C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/trees/kegg.taxa.names.to.ids.txt",sep="\t",stringsAsFactors = F,quote="",header=F)
# colnames(kegg.name.to.taxa.id) = c("name","taxa.id")
# head(kegg.name.to.taxa.id)
# 
# kegg.3letters.to.name =  read.table(file="C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/trees/kegg.taxa.3letters.to.name.txt",sep="\t",stringsAsFactors = F,quote="",header=T)
# kegg.3letters.to.name$name = gsub(" $","",kegg.3letters.to.name$name)
# head(kegg.3letters.to.name)
# dim(kegg.3letters.to.name)
# dim(kegg.name.to.taxa.id)
# head(kegg.name.to.taxa.id)
# head(kegg.name.to.taxa.id$name)
# length(setdiff(kegg.3letters.to.name$name,kegg.name.to.taxa.id$name))
# head(setdiff(kegg.3letters.to.name$name,kegg.name.to.taxa.id$name))
# read.tree(file="C:/xiaoxi/work/papers/SBGNview/package/SBGN-ML.files/data/species.specifid.pathway_pathwayCommons/trees/kegg.commontree.txt")

```
# Check shared UNIPROT protein sets
We think many of the pathwayCommons

# mapped.ids RData
```{r}
library(SBGNview)
data(mapped.ids)
mapped.ids.0=mapped.ids
class(mapped.ids.0)
mapped.ids.0 = as.data.frame(mapped.ids.0,stringsAsFactors=F)
mapped.ids
mapped.ids = list()
mapped.ids$gene = as.vector(t(unique(subset(mapped.ids.0
                                ,cpd.or.gene=="gene"
                                ,select = ID.type.1
                                ))))
mapped.ids$gene = mapped.ids$gene[-which(mapped.ids$gene == "uniref")]
mapped.ids$gene  = c(mapped.ids$gene,"metacyc.SBGN","pathwayCommons")
mapped.ids$gene  = c(mapped.ids$gene,"KO")

mapped.ids$cpd = as.vector(t(unique(subset(mapped.ids.0
                                ,cpd.or.gene=="compound"
                                ,select = ID.type.2
                                ))))
cpd.ids.1 = as.vector(t(unique(subset(mapped.ids.0
                                ,cpd.or.gene=="compound"
                                ,select = ID.type.1
                                ))))
cpd.ids.1
mapped.ids$cpd  = c(mapped.ids$cpd,cpd.ids.1)
mapped.ids$gene = unique(mapped.ids$gene)
mapped.ids$cpd = unique(mapped.ids$cpd)
mapped.ids
save(mapped.ids,file="C:/xiaoxi/work/papers/SBGNview/package/SBGNview/data/mapped.ids.RData")
```








